---
title: "testing file"
format: html
editor: visual
---

# Offset Permanence Review Data Exploration

Here I explore my data set and the patterns in permanence risks across biodiversity and carbon offset programs, policies, temporal trends, geography, project types, ecosystem types, and other core variables.

------------------------------------------------------------------------

## 1. Setup

### 1.1. Load Required Packages

```{r}
#| label: load-packages
#| include: false

# Load all packages 
library(tidyverse) # Data manipulation
library(here) # Easy file paths
library(janitor) # Clean column names
library(readxl) # Reading Excel files
library(countrycode) # Geospatial countries
library(sf) # Geospatial data
library(rnaturalearth) # Geospatial data
library(rnaturalearthdata) # Geospatial data
library(RColorBrewer) # Plot colors
library(knitr) # Knitting document
library(gt) # Tables
library(purrr) # Dataframe manipulation

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### 1.2. Load Helper Functions

```{r}
#| label: load-functions
#| include: false

# Count unique studies mentioning each variable value
summarize_by_study <- function(data, var, label = NULL) {
  var <- rlang::enquo(var)
  label <- label %||% rlang::as_name(var)

  data %>%
    filter(!is.na(!!var), !!var != "") %>%
    group_by(!!var) %>%
    summarise(n_studies = n_distinct(study_title), .groups = "drop") %>%
    arrange(desc(n_studies)) %>%
    rename(!!label := !!var)
}

# Safe combo generator
get_combos <- function(risk_list, combo_size = 2) {
  if (length(risk_list) >= combo_size) {
    combn(risk_list, combo_size, simplify = FALSE)
  } else {
    list()
  }
}

# Count studies with non-missing values
studies_reporting <- function(var) {
  final_df %>%
    filter(!is.na({{ var }})) %>%
    distinct(study_title) %>%
    nrow()
}
```

### 1.3. Load Data

```{r}
#| label: load-data
#| include: false

# Load the cleaned, long, final dataset
final_df <- read_csv(
  here("data", "derived", "offset_perm_rev_long_cleaned.csv"),
  guess_max = 10000,
  col_types = cols(
    species_common_name     = col_character(),
    species_scientific_name = col_character(),
    species_taxonomic_group = col_character(),
    .default = col_guess()
  )
)

# Load risk typology
risk_typology <- read_csv(here("data", "reference", "permanence_risk_typology_lookup.csv")) %>%
  clean_names() %>%
  select(domain = broad, category = specific, type = sub_risk)
```

------------------------------------------------------------------------

## 2. Database Overview

Here I summarize the key characteristics of the final dataset, including ecosystem types, programs, policies, and evidence types.

### 2.1. Key Dataset Statistics

Print basic tables that show the number of studies per category.

```{r}
#| label: key-data-stats
#| echo: false
# Summarize key categorical variables using summarize_by_study()

# How many studies focused on different offset types?
offset_summary    <- summarize_by_study(final_df, offset_category_general)

# What was the distribution of evidence types?
evidence_summary  <- summarize_by_study(final_df, study_evidence_type)

# How many continents and countries were represented?
continent_summary <- summarize_by_study(final_df, continent)
country_summary   <- summarize_by_study(final_df, country)

# How many ecosystems (broad type)?
ecosystem_summary <- summarize_by_study(final_df, ecosystem_broad_type)

# How many specific species ?
species_summary <- summarize_by_study(final_df, species_common_name)

# What sorts of ecological actions were discussed (Broad)?
eco_act_summary  <- summarize_by_study(final_df, project_broad_type)

# Programs and policies
program_summary   <- summarize_by_study(final_df, program_name)
policy_summary    <- summarize_by_study(final_df, policy_name)


# Print summaries
offset_summary
evidence_summary
continent_summary
country_summary  
ecosystem_summary
species_summary
eco_act_summary
program_summary
policy_summary
```

### 2.1.1 Table: Key Dataset Statistics

This table provides an overview of major variables captured in the final dataset, including study characteristics, offset types, ecosystem focus, program and policy mentions, and classifications of permanence risks. For each variable, the table shows the number of unique values observed, how many studies mention the variable, the percentage of total studies, and the top three most frequently cited values (with study counts).

```{r}
#| label: data-summary-table
#| echo: false

# Calculate total unique studies
total_studies <- final_df %>%
  filter(!is.na(study_title), study_title != "") %>%
  distinct(study_title) %>%
  nrow()

# Helper function
summarize_variable <- function(df, var, label, top_n = 3, linebreak = "html") {
  df_clean <- df %>%
    filter(!is.na(.data[[var]]), .data[[var]] != "") %>%
    group_by(study_title) %>%
    summarise(values = list(unique(.data[[var]])), .groups = "drop")

  study_count <- nrow(df_clean)
  all_vals <- unlist(df_clean$values)
  val_counts <- sort(table(all_vals), decreasing = TRUE)
  unique_vals <- length(unique(all_vals))
  top_vals <- head(val_counts, top_n)

  # Choose separator for output format
  sep <- dplyr::case_when(
    linebreak == "html" ~ "<br>",
    linebreak == "newline" ~ "\n",
    TRUE ~ "; "
  )
  top_vals_text <- paste0(names(top_vals), " (n = ", top_vals, ")", collapse = sep)

  tibble(
    Variable = label,
    `Unique Values` = unique_vals,
    `Studies Mentioning` = study_count,
    `% of Total` = paste0(round(100 * study_count / total_studies), "%"),
    `Top 3 Values (N)` = top_vals_text
  )
}

# Variable names and desired row order
var_list <- list(
  c("study_publication_year", "Publication Year"),
  c("offset_category_general", "Offset Type"),
  c("study_evidence_type", "Evidence Type"),
  c("country", "Country"),
  c("ecosystem_broad_type", "Ecosystem (Broad)"),
  c("species_common_name", "Focal Species"),
  c("project_broad_type", "Project Type"),
  c("program_name", "Offset Programs"),
  c("policy_name", "Policies"),
  c("permanence_risk_domain", "Permanence Risk Domains"),
  c("permanence_risk_category", "Permanence Risk Categories"),
  c("permanence_risk_type", "Permanence Risk Types")
)

desired_order <- sapply(var_list, function(x) x[2])

# Generate both versions (HTML and CSV-friendly)
summary_html <- bind_rows(lapply(var_list, function(x) {
  summarize_variable(final_df, var = x[1], label = x[2], linebreak = "html")
})) %>%
  mutate(Variable = factor(Variable, levels = desired_order)) %>%
  arrange(Variable)

summary_csv <- bind_rows(lapply(var_list, function(x) {
  summarize_variable(final_df, var = x[1], label = x[2], linebreak = "newline")
})) %>%
  mutate(Variable = factor(Variable, levels = desired_order)) %>%
  arrange(Variable)

# Display table in Quarto document
summary_html %>%
  gt() %>%
  tab_header(title = "Key Dataset Variables") %>%
  cols_label(
    `Unique Values` = "Unique Values",
    `Studies Mentioning` = "Studies Mentioning",
    `% of Total` = "% of Total",
    `Top 3 Values (N)` = "Top 3 Values (n)"
  ) %>%
  fmt_markdown(columns = vars(`Top 3 Values (N)`))

# Export CSV version to file
write.csv(summary_csv,
          file = here::here("output", "figures", "dataset_sum_table.csv"),
          row.names = FALSE)
```

## 3. Study Characteristics by Offset Category

Let's break down how the key variables differ between **biodiversity** and **carbon** offset studies. For reference, the key variables are:

-   Evidence Type

-   Continent/Country

-   Ecosystem

-   Focal Species

-   Project Type

-   Offset Program

-   Offset Policy

Each figure below shows the number of studies referencing a given value, grouped by offset type.

------------------------------------------------------------------------

### 3.1. Evidence Type x Offset Category

Here I dive into how the evidence types break down across biodiversity and carbon offset studies.Each study was assigned to one of the following evidence types:

1.  Direct Empirical Studies: These assessed permanence risks through primary data collection, either by observing and documenting realized risks (e.g., offset site degradation), or empirically measuring conditions associated with permanence risks (e.g., encroachment, fire frequency). This included field-based assessments, project compliance evaluations, case study analysis, and single- or multi-project site examinations.

2.  Modelling Studies: These evaluated permanence risks using quantitative modelling or simulations rather than direct field observation.

3.  Review and Discussion-based Studies: These synthesized multiple sources to examine permanence risks more broadly and often incorporated both empirical evidence and theoretical insights.

4.  Conceptual, Legal, and Policy-focused Studies: These explored permanence risks from governance, legal, policy, or theoretical perspectives, typically identifying risks through logical reasoning or policy analyses rather than empirical observation .

The table and bar charts below give counts of how many studies fall into each category for each offset type. 
```{r}
#| label: evidence-type-offset-category
#| echo: false

# Create summary table
evidence_offset_summary <- final_df %>%
  filter(
    !is.na(study_evidence_type), study_evidence_type != "",
    !is.na(offset_category_general), offset_category_general != ""
  ) %>%
  group_by(study_evidence_type, offset_category_general) %>%
  summarise(n_studies = n_distinct(study_title), .groups = "drop") %>%
  pivot_wider(names_from = offset_category_general, values_from = n_studies, values_fill = 0) %>%
  mutate(Total = biodiversity + carbon) %>%
  arrange(desc(Total))

# Display table
evidence_offset_summary %>%
  gt() %>%
  tab_header(title = md("**Evidence Types by Offset Category**")) %>%
  cols_label(
    study_evidence_type = "Evidence Type",
    biodiversity = "Biodiversity",
    carbon = "Carbon",
    Total = "Total"
  ) %>%
  fmt_number(columns = c(biodiversity, carbon, Total), decimals = 0) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  )

# Prepare data for plot
evidence_offset_plot <- final_df %>%
  filter(
    !is.na(study_evidence_type), study_evidence_type != "",
    !is.na(offset_category_general), offset_category_general != ""
  ) %>%
  group_by(study_evidence_type, offset_category_general) %>%
  summarise(n_studies = n_distinct(study_title), .groups = "drop")

# Plot
ggplot(evidence_offset_plot,
       aes(x = fct_reorder(study_evidence_type, n_studies, .fun = sum),
           y = n_studies, fill = offset_category_general)) +
  geom_col(position = "dodge", color = "black") +
  geom_text(aes(label = n_studies),
            position = position_dodge(width = 0.9),
            hjust = -0.4, size = 4, color = "black") +
  scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "# Studies by Evidence Type and Offset Category",
    x = "Evidence Type",
    y = "Number of Studies",
    fill = "Offset Category"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    plot.margin = margin(1, 1.5, 1, 1, unit = "lines")
  )
```
The most common evidence type was **`r evidence_offset_summary$study_evidence_type[1]`**
(`r evidence_offset_summary$Total[1]` total studies), with
`r evidence_offset_summary$biodiversity[1]` focused on biodiversity offsets and
`r evidence_offset_summary$carbon[1]` on carbon offsets.

This was followed by **`r evidence_offset_summary$study_evidence_type[2]`**
(`r evidence_offset_summary$Total[2]` studies), then
**`r evidence_offset_summary$study_evidence_type[3]`**
(`r evidence_offset_summary$Total[3]`), and
**`r evidence_offset_summary$study_evidence_type[4]`**
(`r evidence_offset_summary$Total[4]`).

Across all evidence types, the highest number of studies in a single offset category was for
**biodiversity-focused** **`r evidence_offset_summary$study_evidence_type[1]`**
studies (`r evidence_offset_summary$biodiversity[1]` studies).

### 3.2. Geography x Offset Category

```{r}
# Step 1: Country-level data
country_data <- final_df %>%
  filter(
    !is.na(study_title), study_title != "",
    !is.na(country), country != "",
    !is.na(continent), continent != "",
    !is.na(offset_category_general), offset_category_general != ""
  ) %>%
  distinct(study_title, continent, country, offset_category_general) %>%
  count(continent, country, offset_category_general) %>%
  pivot_wider(
    names_from = offset_category_general,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(
    Total = biodiversity + carbon,
    is_total = FALSE
  )

# Step 2: Continent summary rows
continent_data <- country_data %>%
  group_by(continent) %>%
  summarise(
    biodiversity = sum(biodiversity),
    carbon = sum(carbon),
    Total = sum(Total),
    .groups = "drop"
  ) %>%
  mutate(
    country = "Continent Total",
    is_total = TRUE
  )

# Step 3: Combine and arrange
full_table <- bind_rows(country_data, continent_data) %>%  # country rows first
  arrange(continent, is_total, desc(Total))                # sort continent total last

# Step 4: Format with gt
full_table %>%
  gt(rowname_col = "country", groupname_col = "continent") %>%
  tab_header(title = md("**Study Counts by Continent, Country, and Offset Category**")) %>%
  cols_label(
    biodiversity = "Biodiversity",
    carbon = "Carbon",
    Total = "Total"
  ) %>%
  fmt_number(columns = c(biodiversity, carbon, Total), decimals = 0) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_body(rows = is_total),                      # Bold "Total for Continent"
      cells_row_groups()                                # Bold continent headers
    )
  ) %>%
  cols_hide(columns = is_total) %>%                    # Drop the is_total column from view
  cols_width(everything() ~ px(170)) %>%
  tab_options(data_row.padding = px(3))


# create plot

# Prepare data
continent_offset_plot <- final_df %>%
  filter(
    !is.na(study_title),
    !is.na(continent),
    !is.na(offset_category_general)
  ) %>%
  distinct(study_title, continent, offset_category_general) %>%
  count(continent, offset_category_general, name = "n_studies")

# Plot: horizontal bars with same style as evidence_offset_plot
ggplot(continent_offset_plot,
       aes(x = fct_reorder(continent, n_studies, .fun = sum),
           y = n_studies,
           fill = offset_category_general)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  geom_text(aes(label = n_studies),
            position = position_dodge(width = 0.9),
            hjust = -0.4, size = 4, color = "black") +
  scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "# Studies by Continent and Offset Category",
    x = "Continent",
    y = "Number of Studies",
    fill = "Offset Category"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    plot.margin = margin(1, 1.5, 1, 1, unit = "lines")
  )

```
This section highlights how the geography of studies varies by offset type. Overall, `r sum(full_table$Total[full_table$country == "Continent Total"])` studies were distributed across six continents. The highest number came from **`r full_table$continent[which.max(full_table$Total[full_table$country == "Continent Total"])]`**, with `r max(full_table$Total[full_table$country == "Continent Total"])` studies in total. 

In terms of individual countries, the top contributors were:
- **`r full_table$country[which.max(full_table$Total)]`** with `r max(full_table$Total)` total studies,
- followed by **`r full_table$country[order(-full_table$Total)][2]`** (`r full_table$Total[order(-full_table$Total)][2]`),
- and **`r full_table$country[order(-full_table$Total)][3]`** (`r full_table$Total[order(-full_table$Total)][3]`).

Biodiversity offsets were most prevalent in **`r full_table$continent[which.max(full_table$biodiversity[full_table$country == "Continent Total"])]`**, while carbon offsets had stronger representation in **`r full_table$continent[which.max(full_table$carbon[full_table$country == "Continent Total"])]`**.


