---
title: "Publication Figures"
subtitle: "Manuscript Figures for Offset Permanence Review"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
---

# Overview

This document generates all publication-ready figures for the manuscript. All figures use consistent color palettes defined in `code/helpers/color_palettes.R` and are exported as SVG files.

------------------------------------------------------------------------

## Setup

Load required packages, data, and centralized color palettes.

```{r}
#| label: setup
#| include: false

# Source centralized data loading script
source(here::here("code", "helpers", "load_data.R"))

# Source helper functions
source(here::here("code", "helpers", "utils.R"))

# Source centralized color palettes
source(here::here("code", "helpers", "color_palettes.R"))

# Additional packages for specialized figures
library(patchwork)
library(scatterpie)
library(svglite)
library(scales)
```

------------------------------------------------------------------------

# Main Text Figures

## Figure 1: Map of Selected Case Studies

This figure is generated manually in PowerPoint and is not generated in this file. It is a map of the world with the number of studies per country.

## Figure 2: Study Characteristics Overview

This figure is a multi-panel figure summarizing key dataset trends across A) ecosystem type, B) evidence type, C) country, D) program, and E) through time based on publication-year frequency. We generate the individual plots below and combine later.

### Panel A: Ecosystem Types

This panel shows the split of ecosystem types in the dataset. 

```{r}
#| label: summary-panel-a
#| echo: true
#| fig-width: 10
#| fig-height: 5

# Use broad ecosystem types (7 total categories)
ecosystem_df <- final_df |>

  # Filter out missing values
  filter(!is.na(ecosystem_broad_type), !is.na(offset_category_general)) |>
  
  # Remove duplicates
  distinct(study_id, ecosystem_broad_type, offset_category_general) |>
  
  # Count number of studies per ecosystem type
  count(ecosystem_broad_type, offset_category_general, name = "n") |>
  
  # Pivot to wide format
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) |>
  
  # Calculate total number of studies per ecosystem type
  mutate(total = biodiversity + carbon) |>

  # Keep top 5 ecosystem types
  slice_max(total, n = 5, with_ties = FALSE) |>
  
  # Arrange in descending order
  arrange(desc(total)) |>
  
  # Convert ecosystem type to factor
  mutate(ecosystem_broad_type = factor(ecosystem_broad_type, levels = rev(ecosystem_broad_type)))

# Keep long format, filter out zeros so remaining bar takes full width
ecosystem_long <- ecosystem_df |>
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_type", values_to = "n") |>
  filter(n > 0)  # Remove zeros so other category takes full width

# Create position object for consistent use
pos_eco <- position_dodge2(width = 0.9, padding = 0, preserve = "single")

# Grouped bars
panel_ecosystems <- ggplot(ecosystem_long, aes(x = ecosystem_broad_type, y = n, fill = offset_type)) +
  geom_col(position = pos_eco, color = "black", linewidth = 0.3, width = 0.9) +
  geom_text(
    aes(label = n),
    position = pos_eco,
    hjust = -0.3,
    size = 8
  ) +
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(
    x = NULL,
    y = NULL,
    fill = NULL,
    tag = "A"
  ) +
  theme_minimal(base_size = 25) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.tag = element_text(face = "bold", size = 29),
    axis.text.x = element_blank()
  )

print(panel_ecosystems)
```

### Panel B: Evidence Type

The next panel shows the split of evidence types in the dataset.

```{r}
#| label: summary-panel-b-evidence
#| echo: true
#| fig-width: 10
#| fig-height: 5

# Clean up evidence type names
evidence_name_map <- c(
  "direct empirical" = "Direct Empirical Studies",
  "legal and policy" = "Conceptual, Legal & Policy Analyses",
  "review and discussion" = "Reviews & Discussions",
  "modeling" = "Modeling Studies"
)

# Evidence type data with cleaned names
evidence_df <- final_df |>

  # Filter out missing values
  filter(!is.na(study_evidence_type), !is.na(offset_category_general)) |>
  
  # Remove duplicates
  distinct(study_id, study_evidence_type, offset_category_general) |>
  
  # Map evidence type names
  mutate(study_evidence_type = evidence_name_map[study_evidence_type]) |>
  
  # Count number of studies per evidence type
  count(study_evidence_type, offset_category_general, name = "n") |>
  
  # Pivot to wide format
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) |>
  
  # Calculate total number of studies per evidence type
  mutate(total = biodiversity + carbon) |>

  # Arrange in descending order
  arrange(desc(total)) |>
  
  # Convert evidence type to factor
  mutate(study_evidence_type = factor(study_evidence_type, levels = rev(study_evidence_type)))

# Long format, filter out zeros so remaining bar takes full width
evidence_long <- evidence_df |>
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_type", values_to = "n") |>
  filter(n > 0)  # Remove zeros so other category takes full width

# Create position object for consistent use
pos_evi <- position_dodge2(width = 0.9, padding = 0, preserve = "single")

# Create the plot
panel_evidence <- ggplot(evidence_long, aes(x = study_evidence_type, y = n, fill = offset_type)) +
  geom_col(position = pos_evi, color = "black", linewidth = 0.3, width = 0.9) +
  geom_text(
    aes(label = n),
    position = pos_evi,
    hjust = -0.3,
    size = 8
  ) +
  # Use custom colors
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  
  # Adjust y-axis
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  
  # Flip the plot
  coord_flip() +
  
  # Remove axis labels
  labs(
    x = NULL,
    y = NULL, 
    fill = NULL,
    tag = "B"
  ) +
  
  # Use minimal theme
  theme_minimal(base_size = 25) +
  
  # Remove legend, grid, and axis labels
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.tag = element_text(face = "bold", size = 29),
    axis.text.x = element_blank()
  )

print(panel_evidence)
```

### Panel C: Top Countries

This panel shows the top 5 countries with the most studies.

```{r}
#| label: summary-panel-c
#| echo: true
#| fig-width: 10
#| fig-height: 5

# Top countries data
country_df <- final_df |>

  # Filter out missing values
  filter(!is.na(country), !is.na(offset_category_general)) |>
  
  # Remove duplicates
  distinct(study_id, country, offset_category_general) |>
  
  # Recode country names
  mutate(country = recode(country, 
    "United States of America" = "USA", 
    "United Kingdom" = "UK"
  )) |>
  
  # Count number of studies per country
  count(country, offset_category_general, name = "n") |>
  
  # Pivot to wide format
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) |>
  
  # Calculate total number of studies per country
  mutate(total = biodiversity + carbon) |>
  
  # Select top 5 countries
  slice_max(total, n = 5, with_ties = FALSE) |>
  
  # Arrange in descending order
  arrange(desc(total)) |>
  
  # Convert country to factor
  mutate(country = factor(country, levels = rev(country)))

# Keep long format, filter out zeros so remaining bar takes full width
country_long <- country_df |>
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_type", values_to = "n") |>
  filter(n > 0)  # Remove zeros so other category takes full width

# Create position object for consistent use
pos <- position_dodge2(width = 0.9, padding = 0, preserve = "single")

# Create the plot
panel_countries <- ggplot(country_long, aes(x = country, y = n, fill = offset_type)) +
  geom_col(position = pos, color = "black", linewidth = 0.3, width = 0.9) +
  geom_text(
    aes(label = n),
    position = pos,
    hjust = -0.3,
    size = 8
  ) +
  
  # Use custom colors
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  
  # Adjust y-axis
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  
  # Flip the plot
  coord_flip() +
  
  # Remove axis labels
  labs(
    x = NULL,
    y = NULL,
    fill = NULL,
    tag = "C"
  ) +
  
  # Use minimal theme
  theme_minimal(base_size = 25) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.tag = element_text(face = "bold", size = 29),
    axis.text.x = element_blank()
  )

print(panel_countries)
```

### Panel D: Top Offset Programs

This panel shows the top 5 offset programs with the most studies. We group related programs into a single category for display.

```{r}
#| label: summary-panel-d-programs
#| echo: true
#| fig-width: 10
#| fig-height: 6

# Create a program grouping lookup table 
program_group_lookup <- tribble(
  ~program_name,                  ~program_group,                    ~component_label,
  
  # US CWA family
  "US CWA 404 Permitting",        "US Compensatory Mitigation",      NA_character_,
  "US Mitigation Banking",         "US Compensatory Mitigation",      NA_character_,
  
  # California carbon family
  "CA Cap-and-Trade",              "CA ARB Forest Carbon",            NA_character_,
  "California ARB Offset Program", "CA ARB Forest Carbon",            NA_character_,
  "California Forest Carbon Offsets", "CA ARB Forest Carbon",         NA_character_,
  
  # Australia carbon family
  "Australia CFI",                 "AUS Carbon Crediting",            NA_character_,
  "Australia ERF",                 "AUS Carbon Crediting",            NA_character_,
  "Australia ACCU Scheme",         "AUS Carbon Crediting",            NA_character_,
  
  # UNFCCC CDM family
  "AIJ Pilot Phase",               "UNFCCC CDM",                     NA_character_,
  "UNFCCC CDM",                    "UNFCCC CDM",                      NA_character_
)

# Top programs data — use program_group to collapse related programs
program_df <- final_df |>
  
  # Filter out missing values
  filter(!is.na(program_name), !is.na(offset_category_general)) |>
  
  # Left join with program_group_lookup
  left_join(program_group_lookup, by = "program_name") |>
  
  # Coalesce program_group and component_label
  mutate(
    program_group = coalesce(program_group, program_name),
    component_label = coalesce(component_label, NA_character_)
  ) |>
  
  # Remove duplicates
  distinct(study_id, program_group, component_label, offset_category_general) |>
  
  # Count number of studies per program group
  count(program_group, component_label, offset_category_general, name = "n") |>
  
  # Pivot to wide format
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) |>
  
  # Calculate total number of studies per program group
  mutate(total = biodiversity + carbon) |>
  
  # Arrange in descending order
  arrange(desc(total)) |>
  
  # Select top 5 program groups
  slice_head(n = 5) |>
  
  # Add component sub-labels for grouped families
  mutate(
    # Wrap the main group name first to ensure it's clean
    program_group_wrapped = str_wrap(program_group, width = 30),
    display_label = program_group_wrapped
  ) |>
  mutate(display_label = factor(display_label, levels = rev(display_label)))

# Long format, filter out zeros
program_long <- program_df |>
  select(display_label, biodiversity, carbon) |>
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_type", values_to = "n") |>
  filter(n > 0)

# Create position object
pos_prog <- position_dodge2(width = 0.9, padding = 0, preserve = "single")

# Grouped bar plot 
panel_programs <- ggplot(program_long, aes(x = display_label, y = n, fill = offset_type)) +
  
  # Add bars
  geom_col(position = pos_prog, color = "black", linewidth = 0.3, width = 0.9) +
  
  # Add labels
  geom_text(
    aes(label = n),
    position = pos_prog,
    hjust = -0.3,
    size = 8
  ) +
  
  # Use custom colors
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  
  # Adjust y-axis
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  
  # Flip the plot
  coord_flip() +
  
  # Remove axis labels
  labs(
    x = NULL,
    y = NULL,  # Removed "Number of Studies" label
    fill = NULL,
    tag = "D"
  ) +
  
  # Use minimal theme
  theme_minimal(base_size = 25) +
  
  # Remove legend, grid, and axis labels
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.tag = element_text(face = "bold", size = 29),
    axis.text.x = element_blank()
  )

print(panel_programs)
```

#### Quick Check - Program Grouping

Due to how the data was coded, some programs now grouped into families might be double counted. To verify that this is not the case, we compare the original program counts (which would double-count studies tagged with multiple programs) vs. the deduplicated grouped count.

```{r}
#| label: verify-program-grouping
#| echo: true

# Define the four grouped families
grouped_families <- c(
 "US Compensatory Mitigation",
 "CA ARB Forest Carbon",
 "AUS Carbon Crediting",
 "UNFCCC CDM"
)

# --- Table 1: Before vs After counts per group ---
# Individual program counts (naive sum — may double-count)
individual_counts <- final_df |>
 filter(!is.na(program_name), !is.na(offset_category_general)) |>
 left_join(program_group_lookup, by = "program_name") |>
 mutate(program_group = coalesce(program_group, program_name)) |>
 filter(program_group %in% grouped_families) |>
 distinct(study_id, program_name, program_group) |>
 count(program_group, program_name, name = "n_studies_individual")

# Grouped counts (deduplicated — what the figure actually uses)
grouped_counts <- final_df |>
 filter(!is.na(program_name), !is.na(offset_category_general)) |>
 left_join(program_group_lookup, by = "program_name") |>
 mutate(program_group = coalesce(program_group, program_name)) |>
 filter(program_group %in% grouped_families) |>
 distinct(study_id, program_group) |>
 count(program_group, name = "n_studies_grouped")

# Combine for comparison
verification_summary <- individual_counts |>
 group_by(program_group) |>
 summarise(
   member_programs = paste(program_name, " (", n_studies_individual, ")", sep = "", collapse = " + "),
   naive_sum = sum(n_studies_individual),
   .groups = "drop"
 ) |>
 left_join(grouped_counts, by = "program_group") |>
 mutate(
   overlap_studies = naive_sum - n_studies_grouped,
   check = if_else(n_studies_grouped <= naive_sum, "✓ OK", "✗ ERROR")
 )

verification_summary |>
 gt() |>
 cols_label(
   program_group = "Grouped Name",
   member_programs = "Member Programs (individual n)",
   naive_sum = "Naive Sum",
   n_studies_grouped = "Grouped Count",
   overlap_studies = "Overlap (deduplicated)",
   check = "Status"
 ) |>
 tab_header(title = "Program Grouping Verification") |>
 tab_footnote("Grouped Count should always be ≤ Naive Sum. Overlap = studies tagged with multiple programs in the same family.") |>
 opt_stylize(style = 3)
```

And here we check which studies actually map to each group and their original program tags.

```{r}
#| label: verify-study-details
#| echo: true

# --- Table 2: Per-study detail for each group ---
# Shows exactly which studies map to each group and their original program tags
study_detail <- final_df |>
 filter(!is.na(program_name), !is.na(offset_category_general)) |>
 left_join(program_group_lookup, by = "program_name") |>
 mutate(program_group = coalesce(program_group, program_name)) |>
 filter(program_group %in% grouped_families) |>
 distinct(study_id, study_title, program_name, program_group) |>
 group_by(program_group, study_id, study_title) |>
 summarise(
   original_programs = paste(program_name, collapse = " + "),
   n_programs_tagged = n(),
   .groups = "drop"
 ) |>
 arrange(program_group, study_id)

study_detail |>
 gt(groupname_col = "program_group") |>
 cols_label(
   study_id = "Study ID",
   study_title = "Study Title",
   original_programs = "Original Program Tags",
   n_programs_tagged = "# Tags"
 ) |>
 tab_header(title = "Study-Level Detail: Which Studies Map to Each Group") |>
 tab_style(
   style = cell_fill(color = "#FFF3CD"),
   locations = cells_body(rows = n_programs_tagged > 1)
 ) |>
 tab_footnote("Highlighted rows = studies tagged with multiple programs in the same family (these are the ones being deduplicated).") |>
 opt_stylize(style = 3) |>
 cols_width(study_title ~ px(350))
```

### Panel E: Publication Timeline

This panel creates a timeline of studies published per year, each bar has a proportion broken down by offset category.

```{r}
#| label: summary-panel-e
#| echo: true
#| fig-width: 10
#| fig-height: 4

# Timeline data
timeline_df <- final_df |>
  # Filter out missing values 
  filter(!is.na(study_publication_year), !is.na(offset_category_general)) |>
  
  # Remove duplicates
  distinct(study_id, study_publication_year, offset_category_general) |>
  
  # Count number of studies per year and offset category
  count(study_publication_year, offset_category_general, name = "n_studies")

# Calculate totals per year for labels
timeline_totals <- timeline_df |>
  group_by(study_publication_year) |>
  summarise(total = sum(n_studies), .groups = "drop")

# Create the timeline plot
panel_timeline <- ggplot(timeline_df, aes(x = study_publication_year, y = n_studies, fill = offset_category_general)) +
  geom_col(position = "stack", color = "black", linewidth = 0.3) +
  geom_text(
    data = timeline_totals,
    aes(x = study_publication_year, y = total, label = total, fill = NULL),
    vjust = -0.3,
    size = 8,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  scale_x_continuous(breaks = seq(1995, 2025, 5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Publication Year",
    y = NULL,
    fill = NULL,
    tag = "E"
  ) +
  theme_minimal(base_size = 25) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.tag = element_text(face = "bold", size = 29),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.y = element_blank()
  )

print(panel_timeline)
```

### Combined Summary Figure (5 Panels)

Now we combine all panels into a single figure.

```{r}
#| label: summary-combined-5panel
#| echo: true
#| fig-width: 18
#| fig-height: 14

# Layout design:
# Row 1: Evidence (A) | Ecosystems (B)
# Row 2: Countries (C) | Programs (D)
# Row 3: Timeline (E) - full width

# Define explicit layout design
# Each letter represents a panel; duplicate letters to span columns
# Row 1: A|B, Row 2: C|D, Row 3: E (full width)
design <- "
  AABB
  CCDD
  EEEE
"

# Combine all panels using the explicit design
summary_fig <- panel_ecosystems + 
               panel_evidence + 
               panel_countries + 
               panel_programs + 
               panel_timeline +
  plot_layout(design = design, heights = c(1, 1, 0.8)) +
  plot_annotation(
    title = NULL,
    theme = theme(plot.margin = margin(10, 10, 10, 10))
  )

# Remove individual legends, add shared legend
summary_fig <- summary_fig & theme(legend.position = "none")

# Create a version with shared legend
legend_plot <- ggplot(timeline_df, aes(x = study_publication_year, y = n_studies, fill = offset_category_general)) +
  geom_col() +
  scale_fill_manual(values = offset_colors, labels = c("Biodiversity", "Carbon")) +
  theme_minimal() +
  theme(
    legend.position = "top", 
    legend.title = element_blank(),
    legend.text = element_text(size = 27),
    legend.key.size = unit(1.5, "cm")
  )

# Get the legend
library(cowplot)
legend <- get_legend(legend_plot)

# Combine with legend at TOP
final_summary <- plot_grid(
  legend,
  summary_fig,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

print(final_summary)
```

### Save Summary Figure

```{r}
#| label: save-summary-fig
#| echo: false

ggsave(
  filename = here("output", "figures", "main", "fig-2-summary-5-panel.svg"),
  plot = final_summary,
  width = 18,
  height = 20,
  units = "in"
)
```

## Figure 3: Diverging Bar Chart (Relative Risk Prevalence) {#fig-3}

This figure shows the distribution of all permanence risk types across biodiversity and carbon offset studies, normalized by the total number of studies in each category. Bars extend left (percentage of biodiversity studies) and right (percentage of carbon studies), colored by risk category. Dashed lines separate risk categories; solid black lines separate risk domains.

```{r}
#| label: fig-3-diverging-bar-proportions
#| echo: false
#| fig-width: 14
#| fig-height: 12
#| fig-cap: "Relative prevalence of permanence risk types across biodiversity and carbon offset studies. Bars show the percentage of studies in each category reporting a given risk type. Dashed lines separate risk categories; solid lines separate risk domains."

# Get total study counts per offset category
total_studies <- final_df |>
  filter(!is.na(offset_category_general)) |>
  distinct(study_id, offset_category_general) |>
  count(offset_category_general, name = "total_n")

# Prepare and order risk data with proportions
risk_cat_prop <- final_df |>
  filter(
    !is.na(permanence_risk_domain),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_type),
    !is.na(offset_category_general)
  ) |>
  distinct(study_id, permanence_risk_domain, permanence_risk_category, permanence_risk_type,
           offset_category_general) |>
  count(permanence_risk_domain, permanence_risk_category, permanence_risk_type, offset_category_general, name = "n") |>
  left_join(total_studies, by = "offset_category_general") |>
  mutate(prop = (n / total_n) * 100) |>  # Convert to percentage
  select(-n, -total_n) |>
  pivot_wider(names_from = offset_category_general, values_from = prop, values_fill = 0) |>
  mutate(
    total = biodiversity + carbon,
    permanence_risk_domain = factor(permanence_risk_domain, levels = domain_order)
  ) |>
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) |>
  mutate(risk_type_label = factor(permanence_risk_type, levels = rev(unique(permanence_risk_type))))

# Re-order for plotting
risk_cat_prop <- risk_cat_prop |>
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) |>
  mutate(risk_type_label = paste0(permanence_risk_type)) |>
  mutate(risk_type_label = factor(risk_type_label, levels = rev(unique(risk_type_label))))

# Long format for diverging bars
risk_prop_long <- risk_cat_prop |>
  select(permanence_risk_domain, permanence_risk_category, risk_type_label, biodiversity, carbon) |>
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_category", values_to = "prop") |>
  mutate(prop = if_else(offset_category == "biodiversity", -prop, prop))

# Line positions: category-level (reuse from iteration 1)
risk_prop_positions <- risk_cat_prop |>
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) |>
  group_by(permanence_risk_domain, permanence_risk_category) |>
  summarise(max_row = max(row), .groups = "drop") |>
  arrange(permanence_risk_domain, permanence_risk_category) |>
  mutate(line_y = max_row + 0.5)

# Line positions: domain-level
domain_breaks_prop <- risk_cat_prop |>
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) |>
  group_by(permanence_risk_domain) |>
  summarise(max_row = max(row), .groups = "drop") |>
  arrange(permanence_risk_domain) |>
  mutate(line_y = max_row + 0.5)

# Create label dataframe
risk_prop_labels <- risk_prop_long |>
  filter(prop != 0) |>
  mutate(
    hjust_val = if_else(prop < 0, 1.2, -0.2),
    y_val = prop,
    x_val = risk_type_label,
    label_text = paste0(round(abs(prop), 0), "%")
  )

# Apply risk category colors
risk_prop_long <- risk_prop_long |>
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = names(global_risk_colors)))

# Calculate bracket annotation positions
row_positions_prop <- risk_cat_prop |>
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label))))

bracket_data_prop <- row_positions_prop |>
  group_by(permanence_risk_category, permanence_risk_domain) |>
  summarise(
    y_start = min(row),
    y_end = max(row),
    y_mid = (min(row) + max(row)) / 2,
    .groups = "drop"
  ) |>
  mutate(color = global_risk_colors[as.character(permanence_risk_category)])

# Calculate symmetric axis limits for proportions
max_carbon_prop <- max(risk_prop_long$prop[risk_prop_long$prop > 0], na.rm = TRUE)
max_biodiversity_prop <- abs(min(risk_prop_long$prop[risk_prop_long$prop < 0], na.rm = TRUE))
max_abs_prop <- max(max_carbon_prop, max_biodiversity_prop)

# Symmetric limits with buffer
symmetric_limit_prop <- max_abs_prop * 1.5

# Position bracket annotations outside the plot area
bracket_x_prop <- symmetric_limit_prop + 5

# Plot with proportions
p_prop <- ggplot(risk_prop_long, aes(x = risk_type_label, y = prop, fill = permanence_risk_category)) +
  geom_col(color = "black") +
  geom_text(
    data = risk_prop_labels,
    aes(x = x_val, y = y_val, label = label_text, hjust = hjust_val),
    size = 22 / .pt,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1.2) +
  geom_segment(
    data = risk_prop_positions,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "gray70"
  ) +
  geom_segment(
    data = domain_breaks_prop,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  # Bracket vertical lines
  geom_segment(
    data = bracket_data_prop,
    aes(x = y_start - 0.3, xend = y_end + 0.3, y = bracket_x_prop, yend = bracket_x_prop),
    inherit.aes = FALSE,
    color = "gray30",
    linewidth = 0.8
  ) +
  # Bracket top tick
  geom_segment(
    data = bracket_data_prop,
    aes(x = y_start - 0.3, xend = y_start - 0.3, y = bracket_x_prop, yend = bracket_x_prop - 3),
    inherit.aes = FALSE,
    color = "gray30",
    linewidth = 0.8
  ) +
  # Bracket bottom tick
  geom_segment(
    data = bracket_data_prop,
    aes(x = y_end + 0.3, xend = y_end + 0.3, y = bracket_x_prop, yend = bracket_x_prop - 3),
    inherit.aes = FALSE,
    color = "gray30",
    linewidth = 0.8
  ) +
  # Colored squares for categories
  geom_point(
    data = bracket_data_prop,
    aes(x = y_mid, y = bracket_x_prop + 5),
    fill = bracket_data_prop$color,
    color = "black",
    shape = 22,
    size = 9,
    inherit.aes = FALSE
  ) +
  # Category labels
  geom_text(
    data = bracket_data_prop,
    aes(x = y_mid, y = bracket_x_prop + 9, label = permanence_risk_category),
    hjust = 0,
    size = 22 / .pt,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = global_risk_colors, guide = "none") +
  scale_y_continuous(
    labels = function(x) paste0(abs(x), "%"),
    expand = expansion(mult = c(0.02, 0.02)),
    sec.axis = dup_axis(name = NULL, labels = NULL)
  ) +
  coord_flip(clip = "off", ylim = c(-symmetric_limit_prop, symmetric_limit_prop)) +
  labs(
    title = NULL,
    x = NULL,
    y = "Percentage of Studies"
  ) +
  theme_minimal(base_size = 22) +
  theme(
    plot.title = element_blank(),
    axis.text.y = element_text(size = 22),
    axis.text.x = element_text(size = 22),
    axis.title.x.bottom = element_text(size = 22, margin = margin(t = 15)),
    axis.title.x.top = element_text(size = 22, margin = margin(b = 15)),
    axis.title.y = element_text(size = 22, margin = margin(r = 20)),
    legend.position = "none",
    panel.grid = element_blank(),
    plot.margin = margin(t = 1.5, r = 22, b = 1, l = 0.5, unit = "cm")
  )

print(p_prop)
```

### Save Figure 3

```{r}
#| label: save-fig-3-prop
#| echo: false

# SVG (vector)
ggsave(
  filename = here("output", "figures", "main", "fig-3-div-bar-rel-prop.svg"),
  plot = last_plot(),
  width = 24,
  height = 24,
  units = "in"
)
```

## Figure 4: Heatmap of Top Risk Pairs × Offset Category (Domain-Colored)

This figure looks at the 10 most common risk pairs and shows their co-occurrence per offset category.

```{r}
#| label: cooccurrence-heatmap-domain
#| echo: true
#| fig-width: 12
#| fig-height: 10

library(ggtext)

# --- Step 1: Build risk type → category lookup ---
risk_type_info <- final_df |>
  filter(!is.na(permanence_risk_type), !is.na(permanence_risk_category)) |>
  distinct(permanence_risk_type, permanence_risk_category)

# --- Step 2: Generate risk pairs per study × offset type ---
pair_by_offset <- final_df |>
  filter(!is.na(permanence_risk_type), !is.na(offset_category_general)) |>
  distinct(study_id, offset_category_general, permanence_risk_type) |>
  group_by(study_id, offset_category_general) |>
  summarise(risk_pair = list(get_combos(permanence_risk_type, combo_size = 2)), .groups = "drop") |>
  unnest(risk_pair) |>
  mutate(
    risk_a = map_chr(risk_pair, ~ sort(.x)[1]),
    risk_b = map_chr(risk_pair, ~ sort(.x)[2]),
    pair_label = paste0(risk_a, " + ", risk_b)
  ) |>
  count(offset_category_general, pair_label, risk_a, risk_b, name = "n")

# --- Step 3: Top 10 pairs per offset type, then union ---
top_bio <- pair_by_offset |>
  filter(offset_category_general == "biodiversity") |>
  slice_max(n, n = 10, with_ties = FALSE)

top_carbon <- pair_by_offset |>
  filter(offset_category_general == "carbon") |>
  slice_max(n, n = 10, with_ties = FALSE)

all_top_pairs <- unique(c(top_bio$pair_label, top_carbon$pair_label))

# --- Step 4: Filter and complete ---
pair_heat <- pair_by_offset |>
  filter(pair_label %in% all_top_pairs) |>
  complete(offset_category_general, nesting(pair_label, risk_a, risk_b), fill = list(n = 0)) |>
  group_by(pair_label) |>
  mutate(total_n = sum(n)) |>
  ungroup()

# --- Step 5: Build domain-colored labels ---
# Helper to get domain color (same as dot plot)
heat_get_domain_col <- function(risk_name, info_lookup) {
  domain <- info_lookup |>
    filter(permanence_risk_type == risk_name) |>
    pull(permanence_risk_domain) |>
    first()
  case_when(
    domain == "non-physical" ~ "#4E79A7",
    domain == "physical" ~ "#59A14F",
    domain == "methodological" ~ "#E15759",
    TRUE ~ "black"
  )
}

# Build risk-to-domain mapping for the heatmap
risk_info_heat <- final_df |>
  filter(!is.na(permanence_risk_type), !is.na(permanence_risk_domain)) |>
  distinct(permanence_risk_type, permanence_risk_domain)

pair_heat <- pair_heat |>
  mutate(
    colored_label = paste0(
      "<span style='color:", map_chr(risk_a, ~ heat_get_domain_col(.x, risk_info_heat)), "'>", risk_a, "</span> <span style='color:black'>&</span><br>",
      "<span style='color:", map_chr(risk_b, ~ heat_get_domain_col(.x, risk_info_heat)), "'>", risk_b, "</span>"
    )
  )

# --- Step 6: Define custom order (Bio top 10 on top, then Carbon) ---
# Biodiversity top 10 (highest n at the very top)
bio_top_labels <- pair_heat |>
  filter(offset_category_general == "biodiversity", pair_label %in% top_bio$pair_label) |>
  arrange(n) |> 
  pull(colored_label) # Pull the NEW colored labels

# Carbon top 10 excluding any already in bio top 10
carbon_top_labels <- pair_heat |>
  filter(offset_category_general == "carbon", pair_label %in% top_carbon$pair_label) |>
  filter(!(pair_label %in% top_bio$pair_label)) |>
  arrange(n) |>
  pull(colored_label) # Pull the NEW colored labels

label_order <- c(carbon_top_labels, bio_top_labels)

pair_heat <- pair_heat |>
  mutate(colored_label = factor(colored_label, levels = label_order)) |>
  mutate(
    pct = if_else(
      offset_category_general == "biodiversity",
      (n / 80) * 100,
      (n / 57) * 100
    ),
    pct_label = paste0(round(pct, 1), "%")
  )

# --- Step 7: Manual Color Calculation for Faceting ---
# We need to compute the gradient colors manually to use scale_fill_identity
# preventing the need for two separate plots + patchwork
# Bio Gradient: #f3eef8 -> #B19CD9
# Carbon Gradient: #fef9e7 -> #F0C05A


# Calculate distinct max values per market to ensure each reaches full intensity
max_bio_pct <- max(pair_heat$pct[pair_heat$offset_category_general == "biodiversity"], na.rm = TRUE)
max_carbon_pct <- max(pair_heat$pct[pair_heat$offset_category_general == "carbon"], na.rm = TRUE)

# Create color generating functions with market-specific domains
col_fun_bio <- col_numeric(palette = c("#f3eef8", "#B19CD9"), domain = c(0, max_bio_pct))
col_fun_carbon <- col_numeric(palette = c("#fef9e7", "#F0C05A"), domain = c(0, max_carbon_pct))

pair_heat <- pair_heat |>
  mutate(
    fill_color = if_else(
      offset_category_general == "biodiversity",
      col_fun_bio(pct),
      col_fun_carbon(pct)
    ),
    # Ensure text is black (user request)
    text_color = "black" 
  )

# --- Step 8: Single Faceted Plot ---
p_dual_heat <- ggplot(pair_heat, aes(x = "1", y = colored_label, fill = fill_color)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = pct_label), size = 7.5, color = "black") +
  scale_fill_identity() +
  scale_x_discrete(expand = c(0, 0)) +
  facet_grid(. ~ offset_category_general, switch = "x") + # Switch moves strips to bottom (we want top?)
  # Actually, standard facet puts strips on top which acts as headers
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 21) +
  theme(
    axis.text.x = element_blank(), # No x-axis text inside panels, headers are strips
    axis.ticks.x = element_blank(),
    axis.text.y = element_markdown(size = 21, lineheight = 1.3),
    panel.grid = element_blank(),
    # CRITICAL: Eliminate spacing between facets
    panel.spacing = unit(0.1, "lines"), 
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(size = 23, color = "black", margin = margin(b = 10))
  )

# Fix strip labels to be Capitalized
# We can modify the factor levels or use a labeller
label_names <- c("biodiversity" = "Biodiversity", "carbon" = "Carbon")
p_dual_heat <- p_dual_heat + facet_grid(. ~ offset_category_general, labeller = as_labeller(label_names))

print(p_dual_heat)
```

### Save Heatmap

```{r}
#| label: save-cooccurrence-heatmap
#| echo: false

ggsave(
  filename = here("output", "figures", "main", "fig-4-co-occur-heatmap.svg"),
  plot = p_dual_heat,
  width = 11, # Reduced width
  height = 18,
  units = "in"
)
```


## Figure 5: World Map with Risk Distribution {#fig-5}

World map with pie charts showing risk category distribution by continent.

```{r}
#| label: fig-5-world-map-prep
#| echo: false
#| message: false
#| warning: false

# --- Pie data: continent-level risk proportions (used by scatterpie) ---

# Weighted risk category shares per continent
pie_plot_df <- final_df |>
  filter(!is.na(study_id), !is.na(permanence_risk_category),
         !is.na(permanence_risk_domain), !is.na(continent)) |>
  distinct(study_id, continent, permanence_risk_category, permanence_risk_domain) |>
  group_by(study_id, continent) |>
  mutate(weight = 1 / n()) |>
  ungroup() |>
  group_by(continent, permanence_risk_category) |>
  summarise(weighted_sum = sum(weight), .groups = "drop") |>
  group_by(continent) |>
  mutate(
    total_studies = sum(weighted_sum),
    percentage = (weighted_sum / total_studies) * 100
  ) |>
  ungroup() |>
  mutate(permanence_risk_category = factor(permanence_risk_category,
                                           levels = names(global_risk_colors)))

# Store risk category levels for scatterpie cols argument
risk_cat_levels <- levels(pie_plot_df$permanence_risk_category)

# --- Map data: Robinson projection with stadium crop ---

world <- ne_countries(scale = "medium", returnclass = "sf") |>
  filter(sovereignt != "Antarctica")
coast <- ne_download(scale = 50, type = "coastline", category = "physical", returnclass = "sf")

robin_crs <- "+proj=robin"
world_robin <- st_transform(world, crs = robin_crs)
coast_robin <- st_transform(coast, crs = robin_crs)

bbox_crop <- st_bbox(c(xmin = -18000000, xmax = 18000000,
                       ymin = -7000000, ymax = 9000000),
                     crs = robin_crs)
world_crop <- st_intersection(world_robin, st_as_sfc(bbox_crop))
coast_crop <- st_intersection(coast_robin, st_as_sfc(bbox_crop))

# Stadium border shape
make_stadium <- function(xmin, xmax, ymin, ymax, n = 200, crs) {
  r <- (ymax - ymin) / 2
  y_center <- (ymax + ymin) / 2
  theta_left  <- seq(pi / 2, 3 * pi / 2, length.out = n)
  theta_right <- seq(-pi / 2, pi / 2, length.out = n)
  coords <- rbind(
    cbind(xmin + r * cos(theta_left),  y_center + r * sin(theta_left)),
    cbind(xmax + r * cos(theta_right), y_center + r * sin(theta_right))
  )
  coords <- rbind(coords, coords[1, ])
  st_sfc(st_polygon(list(coords)), crs = crs)
}

stadium <- make_stadium(-6000000, 10000000, -7000000, 9000000, crs = robin_crs)
world_stadium <- st_intersection(world_crop, stadium)
coast_stadium <- st_intersection(coast_crop, stadium)

# --- Continent centroids (Robinson) + pie data for scatterpie ---

continent_coords <- data.frame(
  continent = c("Africa", "Asia", "Europe", "North America", "South America", "Oceania"),
  lon = c(25, 90, 10, -100, -60, 140),
  lat = c(5, 40, 52, 40, -15, -30)
)

continent_robin <- st_transform(
  st_as_sf(continent_coords, coords = c("lon", "lat"), crs = 4326),
  crs = robin_crs
)
continent_coords <- cbind(continent_coords, st_coordinates(continent_robin))

continent_abbrev <- c(
  "Africa" = "AF", "Asia" = "AS", "Europe" = "EU",
  "North America" = "NA", "South America" = "SA", "Oceania" = "OC"
)

# Wide-format pie data for scatterpie
pie_df <- pie_plot_df |>
  select(continent, permanence_risk_category, percentage, total_studies) |>
  tidyr::pivot_wider(names_from = permanence_risk_category,
                     values_from = percentage, values_fill = 0) |>
  left_join(continent_coords, by = "continent") |>
  mutate(
    label = paste0(continent_abbrev[continent], " (", round(total_studies), ")"),
    label_X = case_when(
      continent %in% c("South America", "Europe") ~ X - 3e6,
      continent == "Oceania" ~ X - 1e6,
      TRUE ~ X
    ),
    label_Y = case_when(
      continent %in% c("North America", "Africa", "Asia") ~ Y + 2e6,
      continent == "Oceania" ~ Y - 2e6,
      TRUE ~ Y
    )
  )
```

Then plot the figure.

```{r}
#| label: fig-5-world-map
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Global distribution of permanence risk categories. Pie charts show the proportion of risk types mentioned in studies from each continent. Labels show continent abbreviation and total study count."

ggplot() +
  geom_sf(data = world_stadium, fill = "bisque1", color = NA) +
  geom_sf(data = coast_stadium, color = "gray50", size = 0.2) +
  geom_sf(data = stadium, fill = NA, color = "black", size = 0.5) +
  geom_scatterpie(
    data = pie_df,
    aes(x = X, y = Y, r = 1.4e6),
    cols = risk_cat_levels,
    color = "black", size = 0.3
  ) +
  scale_fill_manual(values = global_risk_colors) +
  geom_text(
    data = pie_df,
    aes(x = label_X, y = label_Y, label = label),
    size = 3.5, fontface = "bold"
  ) +
  coord_sf(crs = robin_crs) +
  theme_void(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.margin = margin(t = 5, r = 1, b = 5, l = 1)
  ) +
  guides(fill = guide_legend(nrow = 5, ncol = 2, byrow = TRUE))
```

### Save Figure 5

```{r}
#| label: save-fig-5
#| echo: false

ggsave(
  filename = here("output", "figures", "main", "fig-5-world-map-pies.svg"),
  plot = last_plot(),
  width = 12,
  height = 6,
  units = "in"
)
```

## Figure 6: Temporal Trends by Program {#fig-6}

This figure shows temporal trends in permanence risk categories for the two most common offset programs (REDD+ and U.S. Wetland Mitigation).

```{r}
#| label: fig-6-temporal-programs
#| echo: false
#| fig-width: 12
#| fig-height: 14
#| fig-cap: "Temporal trends in permanence risk categories for the two most common offset programs. Stacked bars show percentage of studies reporting each risk category within 5-year periods. Numbers above bars indicate total studies in each period."

# Filter and bin data
final_df_binned <- final_df %>%
  filter(
    !is.na(study_id),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_domain),
    !is.na(study_publication_year),
    program_name %in% c("REDD+", "US CWA 404 Permitting", "US Mitigation Banking")
  ) %>%
  mutate(
    period = cut(
      study_publication_year,
      breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020, 2026),
      labels = c("1990-1994", "1995-1999", "2000-2004", "2005-2009",
                 "2010-2014", "2015-2019", "2020-2025"),
      right = FALSE
    ),
    program_group = case_when(
      program_name %in% c("US CWA 404 Permitting", "US Mitigation Banking") ~
        "U.S. Compensatory Mitigation",
      TRUE ~ program_name
    )
  )

# Calculate fractional weights per study
study_risk_weights <- final_df_binned %>%
  distinct(study_id, program_group, period, permanence_risk_category) %>%
  group_by(study_id, program_group, period) %>%
  mutate(weight = 1 / n()) %>%
  ungroup()

# Sum weights for proportional representation
risk_shares <- study_risk_weights %>%
  group_by(program_group, period, permanence_risk_category) %>%
  summarise(weighted_sum = sum(weight), .groups = "drop")

# Total studies per period/program
total_study_counts <- final_df_binned %>%
  distinct(study_id, program_group, period) %>%
  count(program_group, period, name = "total_studies")

# Combine and calculate percentages
plot_df <- risk_shares %>%
  left_join(total_study_counts, by = c("program_group", "period")) %>%
  mutate(
    percentage = (weighted_sum / total_studies) * 100,
    label = ifelse(weighted_sum > 0, round(weighted_sum, 1), "")
  )

# Apply global risk colors
plot_df <- plot_df %>%
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = names(global_risk_colors)))

# Plot
ggplot(plot_df, aes(x = period, y = percentage, fill = permanence_risk_category)) +
  geom_col(color = "black") +
  geom_text(
    data = total_study_counts,
    aes(x = period, y = 110, label = total_studies),
    inherit.aes = FALSE, size = 8
  ) +
  facet_wrap(~ program_group, ncol = 1) +
  scale_fill_manual(values = global_risk_colors) +
  labs(
    x = NULL,
    y = "Percentage of Studies",
    fill = "Risk Category"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    strip.text = element_text(size = 20),
    panel.spacing = unit(2, "lines"),
    panel.grid = element_blank(),
    plot.margin = margin(30, 10, 10, 10)
  ) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))
```

### Save Figure 6

```{r}
#| label: save-fig-6
#| echo: false

ggsave(
  filename = here("output", "figures", "main", "fig-6-temporal-risks-program.svg"),
  plot = last_plot(),
  width = 15,
  height = 16,
  units = "in"
)
```