---
title: "extract_figures"
format: html
---

Run this after running the main "permanence risks analysis" file. This is for exporting certain plots for use in the manuscript.

# Plot 1: Diverging Bar Plot (offset type x risk type)

```{r}
#| label: plot-1-diverging-bar-risk-domain-category-type
#| echo: false
#| fig-width: 12
#| fig-height: 10

# Define manual domain order
domain_order <- c("non-physical", "physical", "methodological")

# Apply ordering and arrange data
risk_clean <- final_df %>%
  filter(
    !is.na(permanence_risk_domain),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_type),
    !is.na(offset_category_general)
  ) %>%
  distinct(study_id, permanence_risk_domain, permanence_risk_category, permanence_risk_type,
           offset_category_general) %>%
  count(permanence_risk_domain, permanence_risk_category, permanence_risk_type, offset_category_general, name = "n") %>%
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) %>%
  mutate(
    total = biodiversity + carbon,
    permanence_risk_domain = factor(permanence_risk_domain, levels = domain_order)
  ) %>%
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) %>%
  mutate(risk_type_label = factor(permanence_risk_type, levels = rev(unique(permanence_risk_type))))

# Order for plotting
risk_clean <- risk_clean %>%
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) %>%
  mutate(risk_type_label = paste0(permanence_risk_type)) %>%
  mutate(risk_type_label = factor(risk_type_label, levels = rev(unique(risk_type_label))))

# Long format for diverging bars
risk_long <- risk_clean %>%
  select(permanence_risk_domain, permanence_risk_category, risk_type_label, biodiversity, carbon) %>%
  pivot_longer(cols = c(biodiversity, carbon), names_to = "offset_category", values_to = "n") %>%
  mutate(n = if_else(offset_category == "biodiversity", -n, n))

# Line positions: category-level
risk_positions <- risk_clean %>%
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) %>%
  group_by(permanence_risk_domain, permanence_risk_category) %>%
  summarise(max_row = max(row), .groups = "drop") %>%
  arrange(permanence_risk_domain, permanence_risk_category) %>%
  mutate(line_y = max_row + 0.5)

# Line positions: domain-level
domain_breaks <- risk_clean %>%
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) %>%
  group_by(permanence_risk_domain) %>%
  summarise(max_row = max(row), .groups = "drop") %>%
  arrange(permanence_risk_domain) %>%
  mutate(line_y = max_row + 0.5)

# Create label dataframe with proper coordinates
risk_labels <- risk_long %>%
  filter(n != 0) %>%
  mutate(
    hjust_val = if_else(n < 0, 1.5, -0.5),  # further offset from bars
    y_val = n,
    x_val = risk_type_label
  )

# Plot
ggplot(risk_long, aes(x = risk_type_label, y = n, fill = offset_category)) +
  geom_col(color = "black") +
  geom_text(
    data = risk_labels,
    aes(x = x_val, y = y_val, label = abs(n), hjust = hjust_val),
    size = 7,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_hline(yintercept = 0, color = "gray40") +
  geom_segment(
    data = risk_positions,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "gray70"
  ) +
  geom_segment(
    data = domain_breaks,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  scale_fill_brewer(palette = "Pastel1", name = "Offset Category") +
  scale_y_continuous(
    labels = function(x) abs(x), expand = expansion(mult = c(0.1, 0.15)))+  # adds space on both ends
  coord_flip(clip = "off")+
  labs(
    title = "Risk Type by Offset Category",
    x = NULL,
    y = "Number of Studies"
  ) +
theme_minimal(base_size = 20) +  # base font size
  theme(
    plot.title = element_blank(),
    axis.text.y = element_text(size = 22),                # vertical tick labels (risk types)
    axis.text.x = element_text(size = 22),                # horizontal axis tick labels (values)
    axis.title.x = element_text(size = 24, margin = margin(t = 15)),  # x-axis (bottom, flipped)
    axis.title.y = element_text(size = 24, margin = margin(r = 20)),  # y-axis (left, flipped)
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 22),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "top",
    plot.margin = margin(t = 2, r = 5, b = 2, l = 5, unit = "cm")  # generous padding to avoid cutoffs
  )

```

## Save Plot 1

```{r}
ggsave(
  filename = here("output", "fig-diverging-risk-bar.png"),
  plot = last_plot(),
  width = 20,
  height = 24,
  dpi = 600,
  units = "in"
)
```

## Save Plot 1 (SVG)

```{r}
ggsave(
  filename = here("output", "fig-diverging-risk-bar.svg"),
  plot = last_plot(),
  width = 20,
  height = 24,
  dpi = 600,
  units = "in"
)
```

# Plot 1b: Normalized Diverging Bar Plot (offset type x risk type)

```{r}
#| label: plot-1b-diverging-bar-risk-domain-category-type-normalized
#| echo: false
#| fig-width: 12
#| fig-height: 10

# Use the same data preparation as before
risk_clean_normalized <- final_df %>%
  filter(
    !is.na(permanence_risk_domain),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_type),
    !is.na(offset_category_general)
  ) %>%
  distinct(study_id, permanence_risk_domain, permanence_risk_category, permanence_risk_type,
           offset_category_general) %>%
  count(permanence_risk_domain, permanence_risk_category, permanence_risk_type, offset_category_general, name = "n") %>%
  pivot_wider(names_from = offset_category_general, values_from = n, values_fill = 0) %>%
  mutate(
    total = biodiversity + carbon,
    permanence_risk_domain = factor(permanence_risk_domain, levels = domain_order)
  ) %>%
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) %>%
  mutate(risk_type_label = factor(permanence_risk_type, levels = rev(unique(permanence_risk_type))))

# Order for plotting (same as before)
risk_clean_normalized <- risk_clean_normalized %>%
  arrange(permanence_risk_domain, permanence_risk_category, desc(total)) %>%
  mutate(risk_type_label = paste0(permanence_risk_type)) %>%
  mutate(risk_type_label = factor(risk_type_label, levels = rev(unique(risk_type_label))))

# Calculate max values for normalization by offset category
max_biodiversity <- max(risk_clean_normalized$biodiversity, na.rm = TRUE)
max_carbon <- max(risk_clean_normalized$carbon, na.rm = TRUE)

# Normalize by max values in each category
risk_clean_normalized <- risk_clean_normalized %>%
  mutate(
    biodiversity_norm = biodiversity / max_biodiversity,
    carbon_norm = carbon / max_carbon
  )

# Long format for diverging bars (normalized)
risk_long_normalized <- risk_clean_normalized %>%
  select(permanence_risk_domain, permanence_risk_category, risk_type_label, biodiversity_norm, carbon_norm) %>%
  pivot_longer(cols = c(biodiversity_norm, carbon_norm), names_to = "offset_category", values_to = "n_norm") %>%
  mutate(
    n_norm = if_else(offset_category == "biodiversity_norm", -n_norm, n_norm),
    offset_category = if_else(offset_category == "biodiversity_norm", "biodiversity", "carbon")
  )

# Line positions: category-level (same as before)
risk_positions_normalized <- risk_clean_normalized %>%
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) %>%
  group_by(permanence_risk_domain, permanence_risk_category) %>%
  summarise(max_row = max(row), .groups = "drop") %>%
  arrange(permanence_risk_domain, permanence_risk_category) %>%
  mutate(line_y = max_row + 0.5)

# Line positions: domain-level (same as before)
domain_breaks_normalized <- risk_clean_normalized %>%
  mutate(row = as.numeric(factor(risk_type_label, levels = levels(risk_type_label)))) %>%
  group_by(permanence_risk_domain) %>%
  summarise(max_row = max(row), .groups = "drop") %>%
  arrange(permanence_risk_domain) %>%
  mutate(line_y = max_row + 0.5)

# Create label dataframe with proper coordinates (normalized)
risk_labels_normalized <- risk_long_normalized %>%
  filter(n_norm != 0) %>%
  mutate(
    hjust_val = if_else(n_norm < 0, 1.5, -0.5),  # further offset from bars
    y_val = n_norm,
    x_val = risk_type_label
  )

# Plot (normalized version)
ggplot(risk_long_normalized, aes(x = risk_type_label, y = n_norm, fill = offset_category)) +
  geom_col(color = "black") +
  geom_text(
    data = risk_labels_normalized,
    aes(x = x_val, y = y_val, label = round(abs(n_norm), 2), hjust = hjust_val),
    size = 7,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_hline(yintercept = 0, color = "gray40") +
  geom_segment(
    data = risk_positions_normalized,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "gray70"
  ) +
  geom_segment(
    data = domain_breaks_normalized,
    aes(x = line_y, xend = line_y, y = -Inf, yend = Inf),
    inherit.aes = FALSE,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  scale_fill_brewer(palette = "Pastel1", name = "Offset Category") +
  scale_y_continuous(
    labels = function(x) abs(x), 
    limits = c(-1, 1),
    expand = expansion(mult = c(0.1, 0.15)))+  # adds space on both ends
  coord_flip(clip = "off")+
  labs(
    title = "Risk Type by Offset Category (Normalized by Maximum)",
    x = NULL,
    y = "Normalized Number of Studies"
  ) +
theme_minimal(base_size = 20) +  # base font size
  theme(
    plot.title = element_blank(),
    axis.text.y = element_text(size = 22),                # vertical tick labels (risk types)
    axis.text.x = element_text(size = 22),                # horizontal axis tick labels (values)
    axis.title.x = element_text(size = 24, margin = margin(t = 15)),  # x-axis (bottom, flipped)
    axis.title.y = element_text(size = 24, margin = margin(r = 20)),  # y-axis (left, flipped)
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 22),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "top",
    plot.margin = margin(t = 2, r = 5, b = 2, l = 5, unit = "cm")  # generous padding to avoid cutoffs
  )

```

## Save Plot 1b (Normalized)

```{r}
ggsave(
  filename = here("output", "fig-diverging-risk-bar-normalized.png"),
  plot = last_plot(),
  width = 20,
  height = 24,
  dpi = 600,
  units = "in"
)
```

## Save Plot 1b (Normalized SVG)

```{r}
ggsave(
  filename = here("output", "fig-diverging-risk-bar-normalized.svg"),
  plot = last_plot(),
  width = 20,
  height = 24,
  dpi = 600,
  units = "in"
)
```

# Plot 2: Heatmap of Geographic Continent Trends

```{r}
#| label: risk-heatmap-by-continent-normalized
#| echo: false

# Step 1: Get all combinations of continent and risk category
all_combos <- expand.grid(
  continent = unique(na.omit(final_df$continent)),
  permanence_risk_category = unique(na.omit(final_df$permanence_risk_category)),
  stringsAsFactors = FALSE
)

# Step 2: Count studies by continent × risk category
heatmap_data <- final_df %>%
  filter(!is.na(permanence_risk_category), !is.na(continent)) %>%
  distinct(study_id, continent, permanence_risk_category) %>%
  count(continent, permanence_risk_category, name = "n_studies")

# Step 3: Get total studies per continent
total_per_continent <- final_df %>%
  filter(!is.na(study_id), !is.na(continent)) %>%
  distinct(study_id, continent, permanence_risk_category) %>%
  count(continent, name = "total_studies")

# Step 4: Join and normalize
heatmap_normalized <- all_combos %>%
  left_join(heatmap_data, by = c("continent", "permanence_risk_category")) %>%
  left_join(total_per_continent, by = "continent") %>%
  mutate(
    n_studies = replace_na(n_studies, 0),
    total_studies = replace_na(total_studies, 1),  # Avoid divide-by-zero
    prop = n_studies / total_studies,
    text_color = ifelse(prop <= 0.05, "white", "black"),
    permanence_risk_category = fct_reorder(permanence_risk_category, prop, .fun = sum)
  )

# Optional: Refactor continent order
heatmap_normalized <- heatmap_normalized %>%
  mutate(
    continent = factor(continent, levels = c(
      "North America", "South America", "Europe", "Asia", "Oceania", "Africa"
    ))
  )

# plot

ggplot(heatmap_normalized, aes(x = continent, y = permanence_risk_category, fill = prop)) +
  geom_tile(color = "white") +
  geom_text(
    aes(label = scales::percent(prop, accuracy = 1), color = text_color),
    size = 7
  ) +
  scale_fill_viridis_c(
  trans = "sqrt",  # or "log" for sharper transformation
  option = "viridis",
  limits = c(0, 0.5),
  labels = scales::percent_format(accuracy = 1),
  guide = guide_colorbar(
    barheight = unit(6, "cm"),    # Make legend taller
    barwidth = unit(0.5, "cm"),   # Width of color bar
    title.position = "top",       # Puts title above color bar
    title.hjust = 0.5,             # Center the title
    title.vjust = 4
    )
) +
  scale_color_identity() +
  labs(
    title = NULL,   # Remove title
    x = NULL,       # Remove x-axis label
    y = NULL,       # Remove y-axis label
    fill = "Proportion"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 20),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_blank(),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20)
  )


```

## Save Plot 2

```{r}
ggsave(
  filename = "../output/fig-heatmap-continent.png",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 14,                                          # Width in inches (adjust as needed)
  height = 14,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)

```

# Plot 3: Temporal Permanence Trends by Top 2 Programs

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(colorspace)

# STEP 0: Filter and bin relevant data
final_df_binned <- final_df %>%
  filter(
    !is.na(study_id),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_domain),
    !is.na(study_publication_year),
    program_name %in% c("REDD+", "US CWA 404 Permitting", "US Mitigation Banking")
  ) %>%
  mutate(
    period = cut(
      study_publication_year,
      breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020, 2026),
      labels = c("1990–1994", "1995–1999", "2000–2004", "2005–2009", "2010–2014", "2015–2019", "2020–2025"),
      right = FALSE
    ),
    program_group = case_when(
      program_name %in% c("US CWA 404 Permitting", "US Mitigation Banking") ~ "U.S. CWA Permitting and Mitigation Banking",
      TRUE ~ program_name
    )
  )

# STEP 1: Calculate fractional weights per study
study_risk_weights <- final_df_binned %>%
  distinct(study_id, program_group, period, permanence_risk_category) %>%
  group_by(study_id, program_group, period) %>%
  mutate(weight = 1 / n()) %>%
  ungroup()

# STEP 2: Sum weights to get risk proportions
risk_shares <- study_risk_weights %>%
  group_by(program_group, period, permanence_risk_category) %>%
  summarise(weighted_sum = sum(weight), .groups = "drop")

# STEP 3: Total studies per period/program
total_study_counts <- final_df_binned %>%
  distinct(study_id, program_group, period) %>%
  count(program_group, period, name = "total_studies")

# STEP 4: Combine and calculate %
plot_df <- risk_shares %>%
  left_join(total_study_counts, by = c("program_group", "period")) %>%
  mutate(
    percentage = (weighted_sum / total_studies) * 100,
    label = ifelse(weighted_sum > 0, round(weighted_sum, 1), "")
  )

# STEP 5: Build color palette
all_categories <- final_df %>%
  filter(permanence_risk_category %in% unique(plot_df$permanence_risk_category)) %>%
  distinct(permanence_risk_category, permanence_risk_domain)

domain_base_colors <- c(
  "physical" = "#59A14F",         # green
  "non-physical" = "#4E79A7",     # blue
  "methodological" = "#E15759"    # red
)

get_domain_palette <- function(base_color, n) {
  colorRampPalette(c("#FFFFFF", base_color))(n + 1)[-1]
}

# Assign initial colors
category_colors_df <- all_categories %>%
  group_by(permanence_risk_domain) %>%
  arrange(permanence_risk_category) %>%
  mutate(color = get_domain_palette(domain_base_colors[permanence_risk_domain[1]], n())) %>%
  ungroup()

# STEP 6: Reorder by luminance (dark to light per domain)
rgb_matrix <- hex2RGB(category_colors_df$color)@coords
luminance_vals <- 0.299 * rgb_matrix[, 1] + 0.587 * rgb_matrix[, 2] + 0.114 * rgb_matrix[, 3]

category_order_df <- category_colors_df %>%
  mutate(luminance = luminance_vals) %>%
  mutate(permanence_risk_domain = factor(permanence_risk_domain, levels = c("physical", "non-physical", "methodological"))) %>%
  arrange(permanence_risk_domain, desc(luminance))

# Final ordered levels and named color vector
ordered_levels <- category_order_df$permanence_risk_category
category_colors <- setNames(category_order_df$color, category_order_df$permanence_risk_category)

# Apply to plot_df
plot_df <- plot_df %>%
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = ordered_levels))

# STEP 7: Final Plot
ggplot(plot_df, aes(x = period, y = percentage, fill = permanence_risk_category)) +
  geom_col(color = "black") +
  geom_text(
    data = total_study_counts,
    aes(x = period, y = 105, label = total_studies),
    inherit.aes = FALSE, size = 8
  ) +
  facet_wrap(~ program_group, ncol = 1) +
  scale_fill_manual(values = category_colors) +
  labs(
    x = NULL,
    y = "Percentage of Studies",
    fill = "Risk Category"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    strip.text = element_text(size = 20),
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(30, 10, 10, 10)
  ) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))

```

## Save Plot 3

```{r}
ggsave(
  filename = "../output/fig-temporal-risks-program.png",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 12,                                          # Width in inches (adjust as needed)
  height = 14,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)
```

## Save Plot 3 SVG

```{r}
ggsave(
  filename = "../output/fig-temporal-risks-program.svg",  # Save as SVG
  plot = last_plot(),
  width = 15,
  height = 16,
  units = "in"  # Units still matter for defining physical size
)

```

# Plot 4: Faceted Pie Chart of Permanence Risk Categories by Continent

```{r continent-pie-ordered, fig.width=10, fig.height=6, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(colorspace)

# STEP 0: Filter and prepare data --------------------------------------------
continent_df <- final_df %>%
  filter(
    !is.na(study_id),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_domain),
    !is.na(continent)
  ) %>%
  # Get unique study-category-continent combos
  distinct(study_id, continent, permanence_risk_category, permanence_risk_domain)

# STEP 1: Assign fractional weight per study/continent -----------------------
weighted_df <- continent_df %>%
  group_by(study_id, continent) %>%
  mutate(weight = 1 / n()) %>%
  ungroup()

# STEP 2: Aggregate to get study-weighted counts -----------------------------
continent_risk <- weighted_df %>%
  group_by(continent, permanence_risk_category, permanence_risk_domain) %>%
  summarise(weighted_sum = sum(weight), .groups = "drop")

# STEP 3: Compute percentages per continent ---------------------------------
continent_totals <- continent_risk %>%
  group_by(continent) %>%
  summarise(total_studies = sum(weighted_sum), .groups = "drop")

plot_df <- continent_risk %>%
  left_join(continent_totals, by = "continent") %>%
  mutate(
    percentage = (weighted_sum / total_studies) * 100,
    label = ifelse(percentage >= 5, paste0(round(percentage, 1), "%"), "") # only label slices ≥5%
  )

# STEP 4: Create ordered palette --------------------------------------------
domain_base_colors <- c(
  "physical" = "#59A14F",        # green
  "non-physical" = "#4E79A7",    # blue
  "methodological" = "#E15759"   # red
)

get_domain_palette <- function(base_color, n) {
  colorRampPalette(c("#FFFFFF", base_color))(n + 1)[-1]
}

# Compute base category colors
category_colors_df <- plot_df %>%
  distinct(permanence_risk_category, permanence_risk_domain) %>%
  group_by(permanence_risk_domain) %>%
  arrange(permanence_risk_category) %>%
  mutate(color = get_domain_palette(domain_base_colors[permanence_risk_domain[1]], n())) %>%
  ungroup()

# Compute luminance for dark -> light sorting
rgb_matrix <- hex2RGB(category_colors_df$color)@coords
luminance_vals <- 0.299 * rgb_matrix[, 1] + 0.587 * rgb_matrix[, 2] + 0.114 * rgb_matrix[, 3]

category_colors_df <- category_colors_df %>%
  mutate(luminance = luminance_vals) %>%
  mutate(permanence_risk_domain = factor(permanence_risk_domain, levels = c("physical","non-physical","methodological"))) %>%
  arrange(permanence_risk_domain, luminance)  # dark -> light

# Reorder factor levels and create color vector
ordered_levels <- category_colors_df$permanence_risk_category
category_colors <- setNames(category_colors_df$color, category_colors_df$permanence_risk_category)

plot_df <- plot_df %>%
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = ordered_levels))

# STEP 5: Add facet labels with total studies --------------------------------
facet_labels <- continent_totals %>%
  mutate(facet_label = paste0(continent, " (", round(total_studies), " studies)")) %>%
  select(continent, facet_label)

plot_df <- plot_df %>%
  left_join(facet_labels, by = "continent")

# STEP 6: Plot faceted, ordered pies ----------------------------------------
ggplot(plot_df, aes(x = "", y = percentage, fill = permanence_risk_category)) +
  geom_col(color = "black", size = 0.8, alpha = 0.9)+
  # geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) + # REMOVE THIS
  coord_polar(theta = "y") +
  facet_wrap(~ facet_label) +
  scale_fill_manual(values = category_colors) +
  labs(
    x = NULL, y = NULL,
    fill = "Risk Category",
    title = "Distribution of Permanence Risk Categories by Continent"
  ) +
  theme_void(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    panel.spacing = unit(2, "lines"),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  ) +
  guides(
    fill = guide_legend(
      nrow = 5,
      ncol = 2,
      byrow = TRUE
    )
  )

```

## Save Plot 4

```{r}
ggsave(
  filename = "../output/facet-cont-risk-cat.png",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 12,                                          # Width in inches (adjust as needed)
  height = 14,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)

```

# Plot 4a: Individual Pie Charts (Continent x Risk Category)

```{r}
library(ggplot2)
library(dplyr)

make_pie <- function(df, colors) {
  ggplot(df, aes(x = "", y = percentage, fill = permanence_risk_category)) +
    geom_col(color = "black") +
    coord_polar(theta = "y") +
    scale_fill_manual(values = colors) +
    theme_void() +
    theme(
      legend.position = "none",
      aspect.ratio = 1,  # perfect circle
      plot.margin = margin(0, 0, 0, 0),
      plot.background = element_rect(fill = NA, color = NA),
      panel.background = element_rect(fill = NA, color = NA)
    )
}

facet_list <- unique(plot_df$facet_label)

for (facet in facet_list) {
  df_facet <- plot_df %>% filter(facet_label == facet)
  
  # Extract continent name
  continent_name <- sub(" \\(.*", "", facet)
  file_base <- paste0("pie_", gsub(" ", "_", continent_name))
  
  # Save raw PNG with transparent background
  raw_file <- paste0(file_base, "_raw.png")
  ggsave(
    filename = here("output", raw_file),
    plot = make_pie(df_facet, category_colors),
    width = 6, height = 6, dpi = 600,
    bg = "transparent"
  )
  
}

```

# Plot 4b: Extract the legend

```{r}
library(grid)
library(gtable)

# Add non-breaking spaces to labels for visual padding only
pretty_labels <- paste0(levels(plot_df$permanence_risk_category), "\u00A0\u00A0\u00A0")

legend_plot <- ggplot(plot_df, aes(
  x = permanence_risk_category,
  y = percentage,
  fill = permanence_risk_category
)) +
  geom_col(color = "black") +
  scale_fill_manual(
    values = category_colors,
    labels = pretty_labels   # <--- apply padding here
  ) +
  guides(
    fill = guide_legend(
      nrow = 5,
      ncol = 2,
      byrow = TRUE,
      override.aes = list(color = "black")
    )
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  )


# Convert to gtable
g <- ggplotGrob(legend_plot)

# Locate and extract the guide-box (legend)
legend_index <- which(sapply(g$grobs, function(x) x$name) == "guide-box")
legend <- g$grobs[[legend_index]]

# add spacing
legend$widths <- legend$widths * 1.5

# Optional: Preview
grid.newpage()
grid.draw(legend)

# Save legend as high-res PNG with black outlines
ggsave(
  filename = here::here("output", "risk_category_legend_5x2_black.png"),
  plot = grid::grid.grabExpr(grid::grid.draw(legend)),
  width = 20, height = 4, dpi = 600, bg = "transparent"
)

```

# Plot 5: Nature-style World Map

```{r}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(dplyr)

# 1. Load data
world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(sovereignt != "Antarctica")
coast <- ne_download(scale = 50, type = "coastline", category = "physical", returnclass = "sf")

# 2. Transform to Robinson projection
robin_crs <- "+proj=robin"
world_robin <- st_transform(world, crs = robin_crs)
coast_robin <- st_transform(coast, crs = robin_crs)

# 3. Crop rectangle in projected space
bbox_crop <- st_bbox(c(xmin = -18000000, xmax = 18000000,
                       ymin = -7000000, ymax = 9000000),
                     crs = robin_crs)

world_crop <- st_intersection(world_robin, st_as_sfc(bbox_crop))
coast_crop <- st_intersection(coast_robin, st_as_sfc(bbox_crop))


# 4. Fixed stadium function
make_stadium <- function(xmin, xmax, ymin, ymax, n = 200, crs) {
  r <- (ymax - ymin) / 2  # radius = half of height
  
  # Center y at 0 for symmetry
  y_center <- (ymax + ymin) / 2
  
  # Left semicircle
  theta_left <- seq(pi/2, 3*pi/2, length.out = n)
  left_arc <- cbind(
    xmin + r*cos(theta_left), 
    y_center + r*sin(theta_left)
  )
  
  # Right semicircle
  theta_right <- seq(-pi/2, pi/2, length.out = n)
  right_arc <- cbind(
    xmax + r*cos(theta_right), 
    y_center + r*sin(theta_right)
  )
  
  # Combine into polygon
  coords <- rbind(left_arc, right_arc, left_arc[1, ])
  st_sfc(st_polygon(list(coords)), crs = crs)
}

stadium <- make_stadium(
  xmin = -6000000, xmax = 10000000,  # bring in sides
  ymin = -7000000, ymax = 9000000,
  crs = robin_crs
)

# 5. Clip world and coast to stadium
world_stadium <- st_intersection(world_crop, stadium)
coast_stadium <- st_intersection(coast_crop, stadium)

# 6. Plot Nature-style map
ggplot() +
  geom_sf(data = world_stadium, fill = "bisque1", color = NA) +
  geom_sf(data = coast_stadium, color = "gray50", size = 0.2) +
  geom_sf(data = stadium, fill = NA, color = "black", size = 0.5) +  # outline
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(0,0,0,0)
  )

```

## Save Plot 5

```{r}
ggsave(
  filename = "../output/world_map_nature_style.svg",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 12,                                          # Width in inches (adjust as needed)
  height = 6,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)
```

# Plot 6: World with pie charts

```{r}

# --- 1. Define approximate continent coordinates (lon/lat) ------------------
continent_coords <- data.frame(
  continent = c("Africa", "Asia", "Europe", "North America", "South America", "Oceania"),
  lon = c(25, 90, 10, -100, -60, 140),
  lat = c(5, 40, 52, 40, -15, -30)
)

# --- 2. Project to Robinson CRS ---
continent_sf <- st_as_sf(continent_coords, coords = c("lon", "lat"), crs = 4326)
continent_robin <- st_transform(continent_sf, crs = robin_crs)
continent_coords_robin <- st_coordinates(continent_robin)
continent_coords <- cbind(continent_coords, continent_coords_robin)

# --- 3. Create abbreviation + study count labels ---
continent_abbrev <- c(
  "Africa" = "AF",
  "Asia" = "AS",
  "Europe" = "EU",
  "North America" = "NA",
  "South America" = "SA",
  "Oceania" = "OC"
)

# Join totals for labeling
pie_df <- plot_df %>%
  select(continent, permanence_risk_category, percentage, total_studies) %>%
  tidyr::pivot_wider(names_from = permanence_risk_category, 
                     values_from = percentage, values_fill = 0) %>%
  left_join(continent_coords, by = "continent") %>%
  mutate(
    abbrev = continent_abbrev[continent],
    label = paste0(abbrev, " (", round(total_studies), ")")
  )

# --- 4. Add manual label offsets -------------------------------------------
pie_df <- pie_df %>%
  mutate(
    label_X = X,  # start at pie center
    label_Y = Y
  ) %>%
  mutate(
    label_X = case_when(
      continent == "North America" ~ X,         # no horizontal move
      continent == "South America" ~ X - 3e6,   # move left
      continent == "Africa"        ~ X,         # no horizontal move
      continent == "Europe"        ~ X - 3e6,   # move left
      continent == "Oceania"       ~ X - 1e6,   # move left
      continent == "Asia"          ~ X,         # no horizontal move
      TRUE ~ X
    ),
    label_Y = case_when(
      continent == "North America" ~ Y + 2e6,   # move up
      continent == "South America" ~ Y,         # no vertical move
      continent == "Africa"        ~ Y + 2e6,   # move up
      continent == "Europe"        ~ Y,         # no vertical move
      continent == "Oceania"       ~ Y - 2e6,   # move down
      continent == "Asia"          ~ Y + 2e6,   # move up
      TRUE ~ Y
    )
  )

# --- 5. Plot map + pies with manually moved labels -------------------------
ggplot() +
  # Base map
  geom_sf(data = world_stadium, fill = "bisque1", color = NA) +
  geom_sf(data = coast_stadium, color = "gray50", size = 0.2) +
  geom_sf(data = stadium, fill = NA, color = "black", size = 0.5) +
  
  # Pies
  geom_scatterpie(
    data = pie_df,
    aes(x = X, y = Y, r = 1.4e6),  # adjust radius
    cols = levels(plot_df$permanence_risk_category),
    color = "black", size = 0.3
  ) +
  scale_fill_manual(values = category_colors) +
  
  # Labels at adjusted positions
  geom_text(
    data = pie_df,
    aes(x = label_X, y = label_Y, label = label),
    size = 3.5,
    fontface = "bold"
  ) +
  
  coord_sf(crs = robin_crs) +
  theme_void(base_size = 16) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.margin = margin(t = 5, r = 1, b = 5, l = 1) # extra bottom space
  ) +
  guides(
    fill = guide_legend(nrow = 5, ncol = 2, byrow = TRUE)
  )

```

## Save Plot 6

```{r}
ggsave(
  filename = "../output/world_map_pie_charts_leg_top.svg",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 12,                                          # Width in inches (adjust as needed)
  height = 6,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)
```

# Plot 7: Individual continent maps

```{r}
# Packages
library(sf)
library(ggplot2)
library(rnaturalearth)
library(dplyr)

# 1. Load world map
world <- ne_countries(scale = "medium", returnclass = "sf")

# ---- 2. Define continent membership manually ----
continent_countries <- list(
  
  "North America" = c(
    "Canada", "United States of America", "Mexico", "Greenland",
    "Guatemala", "Belize", "Honduras", "El Salvador", "Nicaragua", 
    "Costa Rica", "Panama", "The Bahamas", "Cuba", "Jamaica", 
    "Haiti", "Dominican Republic", "Trinidad and Tobago", "Barbados",
    "Saint Lucia", "Saint Vincent and the Grenadines", "Grenada",
    "Saint Kitts and Nevis", "Antigua and Barbuda", "Dominica"
  ),
  
  "South America" = c(
    "Colombia", "Venezuela", "Guyana", "Suriname", "Ecuador",
    "Peru", "Brazil", "Bolivia", "Paraguay", "Chile", 
    "Argentina", "Uruguay", "Falkland Islands", "French Guiana"
  ),
  
  "Europe" = c(
    "Portugal", "Spain", "France", "Belgium", "Netherlands", "Germany",
    "Luxembourg", "Switzerland", "Austria", "Italy", "Slovenia", "Croatia",
    "Bosnia and Herzegovina", "Montenegro", "Albania", "Greece", 
    "North Macedonia", "Bulgaria", "Romania", "Moldova", 
    "Hungary", "Slovakia", "Czechia", "Poland", "Lithuania", 
    "Latvia", "Estonia", "Finland", "Sweden", "Norway", "Denmark", 
    "Iceland", "Ireland", "United Kingdom", "Belarus", "Ukraine",
    "Turkey", "Russia", "Georgia", "Armenia", "Azerbaijan",
    "Republic of Serbia", "Malta", "Liechtenstein", "Andorra", 
    "Monaco", "San Marino", "Kosovo", "Vatican"
  ),
  
  "Africa" = c(
    "Morocco", "Algeria", "Tunisia", "Libya", "Egypt", "Mauritania", 
    "Mali", "Senegal", "Gambia", "Guinea-Bissau", "Guinea", "Sierra Leone",
    "Liberia", "Ivory Coast", "Ghana", "Togo", "Benin", "Burkina Faso", 
    "Niger", "Nigeria", "Chad", "Sudan", "South Sudan", "Eritrea", 
    "Djibouti", "Somalia", "Ethiopia", "Central African Republic", "Cameroon",
    "Equatorial Guinea", "Gabon", "Republic of the Congo", "Democratic Republic of the Congo",
    "Uganda", "Kenya", "Rwanda", "Burundi", "United Republic of Tanzania",
    "Angola", "Zambia", "Malawi", "Mozambique", "Zimbabwe", "Namibia", 
    "Botswana", "South Africa", "Lesotho", "eSwatini", "Madagascar", 
    "Cabo Verde", "São Tomé and Principe", "Comoros", "Somaliland", "Western Sahara"
  ),
  
  "Asia" = c(
    "China", "India", "Pakistan", "Afghanistan", "Nepal", "Bhutan", 
    "Bangladesh", "Myanmar", "Thailand", "Laos", "Cambodia", "Vietnam", 
    "Malaysia", "Singapore", "Indonesia", "Philippines", "Brunei", 
    "Mongolia", "Japan", "North Korea", "South Korea", 
    "Kazakhstan", "Uzbekistan", "Turkmenistan", "Kyrgyzstan", "Tajikistan", 
    "Iran", "Iraq", "Syria", "Lebanon", "Israel", "Jordan", 
    "Saudi Arabia", "Yemen", "Oman", "United Arab Emirates", "Qatar", 
    "Bahrain", "Kuwait", "Georgia", "Armenia", "Azerbaijan", 
    "Russia", "Turkey", "East Timor", "Taiwan", "Sri Lanka", "Cyprus"
  ),
  
  "Oceania" = c(
    "Australia", "New Zealand", "Papua New Guinea", "Fiji", "Solomon Islands", 
    "Vanuatu", "Samoa", "Tonga", "Kiribati", "Tuvalu", "Palau", 
    "Marshall Islands", "Nauru", "Federated States of Micronesia"
  )
)

# ---- 3. Create sf object for each continent ----
continent_sf <- lapply(names(continent_countries), function(cont) {
  world %>% filter(admin %in% continent_countries[[cont]])
})
names(continent_sf) <- names(continent_countries)

# ---- 4. Manually expanded bounding boxes (~20% bigger) ----
continent_bboxes <- list(
  "North America" = c(xmin = -180, xmax = -20, ymin = -5, ymax = 90),
  "South America" = c(xmin = -110, xmax = -15, ymin = -70, ymax = 25),
  "Europe"        = c(xmin = -35, xmax = 65, ymin = 25, ymax = 80),
  "Africa"        = c(xmin = -35, xmax = 70, ymin = -50, ymax = 45),
  "Asia"          = c(xmin = 15, xmax = 190, ymin = -25, ymax = 90),
  "Oceania"       = c(xmin = 90, xmax = 190, ymin = -60, ymax = 15)
)

# ---- 5. Plotting function ----
plot_continent <- function(continent_name, bbox_coords) {
  bbox <- st_bbox(bbox_coords, crs = st_crs(4326)) %>% st_as_sfc()
  
  ggplot() +
    geom_sf(data = continent_sf[[continent_name]], 
            fill = "bisque1", color = "gray40", size = 0.2) +
    geom_sf(data = bbox, fill = NA, color = "red", linetype = "dashed") +
    coord_sf(
      xlim = c(bbox_coords["xmin"], bbox_coords["xmax"]),
      ylim = c(bbox_coords["ymin"], bbox_coords["ymax"]),
      expand = FALSE
    ) +
    theme_void(base_size = 14)
}

# ---- 6. Generate plots ----
plots <- lapply(names(continent_bboxes), function(cont) {
  plot_continent(cont, continent_bboxes[[cont]])
})

# Example: view them
# plots[[1]]  # North America
# plots[[2]]  # South America
# plots[[3]]  # Europe
# plots[[4]]  # Africa
# plots[[5]]  # Asia
# plots[[6]]  # Oceania


# ---- Clean plotting function (no red dashed box) ----
plot_continent_clean <- function(continent_name, bbox_coords) {
  ggplot() +
    geom_sf(data = continent_sf[[continent_name]], 
            fill = "bisque1", color = "gray40", size = 0.2) +
    coord_sf(
      xlim = c(bbox_coords["xmin"], bbox_coords["xmax"]),
      ylim = c(bbox_coords["ymin"], bbox_coords["ymax"]),
      expand = FALSE
    ) +
    theme_void(base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold")
    )
}

# ---- Generate clean plots ----
clean_plots <- lapply(names(continent_bboxes), function(cont) {
  plot_continent_clean(cont, continent_bboxes[[cont]])
})
names(clean_plots) <- names(continent_bboxes)


# Example: view them
clean_plots[[1]]  # North America
clean_plots[[2]]  # South America
clean_plots[[3]]  # Europe
clean_plots[[4]]  # Africa
clean_plots[[5]]  # Asia
clean_plots[[6]]  # Oceania
```

## Save Plot 7

```{r}
library(svglite)

# ---- Export each plot as SVG ----
output_dir <- "output"    # choose your folder
dir.create(output_dir, showWarnings = FALSE)

for (i in seq_along(clean_plots)) {
  continent_name <- names(clean_plots)[i]
  file_path <- file.path(output_dir, paste0(continent_name, ".svg"))
  
  # Force independent svglite device
  svglite::svglite(file_path, width = 10, height = 10)
  print(clean_plots[[i]])
  dev.off()
}
```

## Save Plot 7

```{r}
library(svglite)

# ---- Export each plot as SVG ----
output_dir <- "output"    # choose your folder
dir.create(output_dir, showWarnings = FALSE)

for (i in seq_along(clean_plots)) {
  continent_name <- names(clean_plots)[i]
  file_path <- file.path(output_dir, paste0(continent_name, ".svg"))
  
  # Force independent svglite device
  svglite::svglite(file_path, width = 10, height = 10)
  print(clean_plots[[i]])
  dev.off()
}
```

# Plot 3b: Temporal Permanence Trends by Top 2 Programs (Matching Colors)

```{r}
 library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(colorspace)

# STEP 0: Filter and bin relevant data (same as original)
final_df_binned <- final_df %>%
  filter(
    !is.na(study_id),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_domain),
    !is.na(study_publication_year),
    program_name %in% c("REDD+", "US CWA 404 Permitting", "US Mitigation Banking")
  ) %>%
  mutate(
    period = cut(
      study_publication_year,
      breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020, 2026),
      labels = c("1990–1994", "1995–1999", "2000–2004", "2005–2009", "2010–2014", "2015–2019", "2020–2025"),
      right = FALSE
    ),
    program_group = case_when(
      program_name %in% c("US CWA 404 Permitting", "US Mitigation Banking") ~ "U.S. CWA Permitting and Mitigation Banking",
      TRUE ~ program_name
    )
  )

# STEP 1: Calculate fractional weights per study (same as original)
study_risk_weights <- final_df_binned %>%
  distinct(study_id, program_group, period, permanence_risk_category) %>%
  group_by(study_id, program_group, period) %>%
  mutate(weight = 1 / n()) %>%
  ungroup()

# STEP 2: Sum weights to get risk proportions (same as original)
risk_shares <- study_risk_weights %>%
  group_by(program_group, period, permanence_risk_category) %>%
  summarise(weighted_sum = sum(weight), .groups = "drop")

# STEP 3: Total studies per period/program (same as original)
total_study_counts <- final_df_binned %>%
  distinct(study_id, program_group, period) %>%
  count(program_group, period, name = "total_studies")

# STEP 4: Combine and calculate % (same as original)
plot_df <- risk_shares %>%
  left_join(total_study_counts, by = c("program_group", "period")) %>%
  mutate(
    percentage = (weighted_sum / total_studies) * 100,
    label = ifelse(weighted_sum > 0, round(weighted_sum, 1), "")
  )

# STEP 5: Build CONSISTENT color palette (same as Plot 4)
all_categories <- final_df %>%
  filter(permanence_risk_category %in% unique(plot_df$permanence_risk_category)) %>%
  distinct(permanence_risk_category, permanence_risk_domain)

domain_base_colors <- c(
  "physical" = "#59A14F",         # green
  "non-physical" = "#4E79A7",     # blue
  "methodological" = "#E15759"    # red
)

get_domain_palette <- function(base_color, n) {
  colorRampPalette(c("#FFFFFF", base_color))(n + 1)[-1]
}

# Assign initial colors
category_colors_df <- all_categories %>%
  group_by(permanence_risk_domain) %>%
  arrange(permanence_risk_category) %>%
  mutate(color = get_domain_palette(domain_base_colors[permanence_risk_domain[1]], n())) %>%
  ungroup()

# STEP 6: Reorder by luminance (dark to light per domain) - SAME AS PLOT 4
rgb_matrix <- hex2RGB(category_colors_df$color)@coords
luminance_vals <- 0.299 * rgb_matrix[, 1] + 0.587 * rgb_matrix[, 2] + 0.114 * rgb_matrix[, 3]

category_order_df <- category_colors_df %>%
  mutate(luminance = luminance_vals) %>%
  mutate(permanence_risk_domain = factor(permanence_risk_domain, levels = c("physical", "non-physical", "methodological"))) %>%
  arrange(permanence_risk_domain, desc(luminance))

# Final ordered levels and named color vector
ordered_levels <- category_order_df$permanence_risk_category
category_colors <- setNames(category_order_df$color, category_order_df$permanence_risk_category)

# Apply to plot_df
plot_df <- plot_df %>%
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = ordered_levels))

# STEP 7: Final Plot with CONSISTENT colors
ggplot(plot_df, aes(x = period, y = percentage, fill = permanence_risk_category)) +
  geom_col(color = "black") +
  geom_text(
    data = total_study_counts,
    aes(x = period, y = 110, label = total_studies),
    inherit.aes = FALSE, size = 8
  ) +
  facet_wrap(~ program_group, ncol = 1) +
  scale_fill_manual(values = category_colors) +
  labs(
    x = NULL,
    y = "Percentage of Studies",
    fill = "Risk Category"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    strip.text = element_text(size = 20),
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(30, 10, 10, 10)
  ) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))

```

## Save Plot 3b (Matching Colors)



```{r}
ggsave(
  filename = "../output/fig-temporal-risks-program-matching-colors.png",  # Save to output folder
  plot = last_plot(),                                 # Or specify your plot object
  width = 12,                                          # Width in inches (adjust as needed)
  height = 14,                                         # Height in inches (adjust as needed)
  dpi = 600,                                          # High-resolution
  units = "in"                                        # Use inches for journals
)
```

## Save Plot 3b (Matching Colors SVG)



```{r}
ggsave(
  filename = "../output/fig-temporal-risks-program-matching-colors.svg",  # Save as SVG
  plot = last_plot(),
  width = 15,
  height = 16,
  units = "in"  # Units still matter for defining physical size
)
```


# Plot 6b: World with pie charts (Matching Colors)

```{r}
library(scatterpie)

# Use the same continent data from Plot 4
continent_df <- final_df %>%
  filter(
    !is.na(study_id),
    !is.na(permanence_risk_category),
    !is.na(permanence_risk_domain),
    !is.na(continent)
  ) %>%
  distinct(study_id, continent, permanence_risk_category, permanence_risk_domain)

# Assign fractional weight per study/continent (same as Plot 4)
weighted_df <- continent_df %>%
  group_by(study_id, continent) %>%
  mutate(weight = 1 / n()) %>%
  ungroup()

# Aggregate to get study-weighted counts (same as Plot 4)
continent_risk <- weighted_df %>%
  group_by(continent, permanence_risk_category, permanence_risk_domain) %>%
  summarise(weighted_sum = sum(weight), .groups = "drop")

# Compute percentages per continent (same as Plot 4)
continent_totals <- continent_risk %>%
  group_by(continent) %>%
  summarise(total_studies = sum(weighted_sum), .groups = "drop")

plot_df <- continent_risk %>%
  left_join(continent_totals, by = "continent") %>%
  mutate(
    percentage = (weighted_sum / total_studies) * 100,
    label = ifelse(percentage >= 5, paste0(round(percentage, 1), "%"), "") # only label slices ≥5%
  )

# Use the SAME color palette as Plot 4 and Plot 3b
plot_df <- plot_df %>%
  mutate(permanence_risk_category = factor(permanence_risk_category, levels = ordered_levels))

# --- 1. Define approximate continent coordinates (lon/lat) ------------------
continent_coords <- data.frame(
  continent = c("Africa", "Asia", "Europe", "North America", "South America", "Oceania"),
  lon = c(25, 90, 10, -100, -60, 140),
  lat = c(5, 40, 52, 40, -15, -30)
)

# --- 2. Project to Robinson CRS ---
continent_sf <- st_as_sf(continent_coords, coords = c("lon", "lat"), crs = 4326)
continent_robin <- st_transform(continent_sf, crs = robin_crs)
continent_coords_robin <- st_coordinates(continent_robin)
continent_coords <- cbind(continent_coords, continent_coords_robin)

# --- 3. Create abbreviation + study count labels ---
continent_abbrev <- c(
  "Africa" = "AF",
  "Asia" = "AS",
  "Europe" = "EU",
  "North America" = "NA",
  "South America" = "SA",
  "Oceania" = "OC"
)

# Join totals for labeling
pie_df <- plot_df %>%
  select(continent, permanence_risk_category, percentage, total_studies) %>%
  tidyr::pivot_wider(names_from = permanence_risk_category, 
                     values_from = percentage, values_fill = 0) %>%
  left_join(continent_coords, by = "continent") %>%
  mutate(
    abbrev = continent_abbrev[continent],
    label = paste0(abbrev, " (", round(total_studies), ")")
  )

# --- 4. Add manual label offsets -------------------------------------------
pie_df <- pie_df %>%
  mutate(
    label_X = X,  # start at pie center
    label_Y = Y
  ) %>%
  mutate(
    label_X = case_when(
      continent == "North America" ~ X,         # no horizontal move
      continent == "South America" ~ X - 3e6,   # move left
      continent == "Africa"        ~ X,         # no horizontal move
      continent == "Europe"        ~ X - 3e6,   # move left
      continent == "Oceania"       ~ X - 1e6,   # move left
      continent == "Asia"          ~ X,         # no horizontal move
      TRUE ~ X
    ),
    label_Y = case_when(
      continent == "North America" ~ Y + 2e6,   # move up
      continent == "South America" ~ Y,         # no vertical move
      continent == "Africa"        ~ Y + 2e6,   # move up
      continent == "Europe"        ~ Y,         # no vertical move
      continent == "Oceania"       ~ Y - 2e6,   # move down
      continent == "Asia"          ~ Y + 2e6,   # move up
      TRUE ~ Y
    )
  )

# --- 5. Plot map + pies with CONSISTENT colors -------------------------
ggplot() +
  # Base map
  geom_sf(data = world_stadium, fill = "bisque1", color = NA) +
  geom_sf(data = coast_stadium, color = "gray50", size = 0.2) +
  geom_sf(data = stadium, fill = NA, color = "black", size = 0.5) +
  
  # Pies with CONSISTENT colors
  geom_scatterpie(
    data = pie_df,
    aes(x = X, y = Y, r = 1.4e6),  # adjust radius
    cols = levels(plot_df$permanence_risk_category),
    color = "black", size = 0.3
  ) +
  scale_fill_manual(values = category_colors) +
  
  # Labels at adjusted positions
  geom_text(
    data = pie_df,
    aes(x = label_X, y = label_Y, label = label),
    size = 3.5,
    fontface = "bold"
  ) +
  
  coord_sf(crs = robin_crs) +
  theme_void(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.margin = margin(t = 5, r = 1, b = 5, l = 1) # extra bottom space
  ) +
  guides(
    fill = guide_legend(nrow = 5, ncol = 2, byrow = TRUE)
  )

```

## Save Plot 6b (Matching Colors)



```{r}

output_file <- "output/world_map_pie_charts_leg_top-matching-colors.svg" 

ggsave(
  filename = output_file,
  plot = last_plot(), 
  width = 12, 
  height = 6, 
  dpi = 600, 
  units = "in" 
)
